---
title: "VenAza_Feb27"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M')`"
output:
  pdf_document:
    latex_engine: lualatex
    keep_tex: true
    df_print: tibble
  html_document: default
---

## Load Library
```{r, eval=FALSE, echo =FALSE, message=FALSE, warning=FALSE}
library(rpx)
#library(msdata)
library(Spectra)
library(tidyverse)
#library(cleaver)
library(MSnbase)
library(tidyverse)
library(mzR)
library(PSMatch)
library(pheatmap)
library(QFeatures)
#library(sageR)
library(SummarizedExperiment)
library(factoextra)
library(limma)
library(DEP)
library(ggrepel)
library(data.table)

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("rpx")
```

## 1. Read evidence.txt
```{r}
library(data.table)
## Read raw peptide TSV
evidence <- read_tsv("/Users/nvd448/Documents/VenAza Project/evidence.txt", show_col_types = FALSE)

setDT(evidence)
```

## 2. Identify TMT reporter intensity columns and Mapping metadata
```{r}
evidence <- evidence[(is.na(Reverse) | Reverse != "+") & (is.na(`Potential contaminant`) | `Potential contaminant` != "+")]
tmt_cols <- grep("^Reporter intensity corrected", colnames(evidence), value = TRUE)
evidence <- evidence[rowSums(as.matrix(evidence[, tmt_cols, with = FALSE])) > 0, ]

library (readxl)
map <- read_excel("~/Documents/VenAza Project/5LABELS_CoPISA_E4.xlsx", sheet = "Sheet1" )

```

## 3. Build assay matrices
```{r}

library(data.table)
library(dplyr)
library(stringr)
library(tibble)
library(QFeatures)
library(SummarizedExperiment)

# ---------------------
# Step 1: Filter evidence
# ---------------------
evidence <- evidence[(is.na(Reverse) | Reverse != "+") &
                     (is.na(`Potential contaminant`) | `Potential contaminant` != "+")]

tmt_cols <- grep("^Reporter intensity corrected", colnames(evidence), value = TRUE)

evidence <- evidence[
  rowSums(as.matrix(evidence[, tmt_cols, with = FALSE])) > 0
]

# ---------------------
# Step 2: Prepare sample annotation
# ---------------------
sample_annotation <- map %>%
  select(label, treatment, cell, `repeat`, sample, Experiment) %>%
  distinct() %>%
  filter(`Experiment` != 4) %>% 
  mutate(
    TMT_channel = str_replace(label, "TMTpro-", "TMT"),
    TMT_channel = gsub(" ", "", TMT_channel),
    Experiment = paste0("Kit", Experiment)
  )

# ---------------------
# Step 3: Define TMTpro 16plex channel order
# ---------------------
channel_map <- c(
  "TMT126","TMT127N","TMT127C","TMT128N",
  "TMT128C","TMT129N","TMT129C","TMT130N",
  "TMT130C","TMT131N","TMT131C","TMT132N",
  "TMT132C","TMT133N","TMT133C","TMT134N"
)

# ---------------------
# Step 4: Build QFeatures object
# ---------------------
qf <- QFeatures()

experiments <- unique(evidence$Experiment)

for (exp in experiments) {
  
  message("Processing: ", exp)
  
  exp_data <- evidence[Experiment == exp]
  
  tmt_cols_exp <- grep("^Reporter intensity corrected",
                       colnames(exp_data),
                       value = TRUE)
  
  # Safety check
  stopifnot(length(tmt_cols_exp) == 16)
  
  # ---- Match annotation to channels ----
  used_channels <- sample_annotation %>%
    filter(Experiment == exp) %>%
    pull(TMT_channel)
  
  keep_idx <- match(used_channels, channel_map)
  keep_idx <- keep_idx[!is.na(keep_idx)]
  
  # ---- Subset intensity columns ----
  exp_subset <- exp_data[, tmt_cols_exp[keep_idx], with = FALSE]
  
  mat <- as.matrix(exp_subset)
  mode(mat) <- "numeric"
  
  colnames(mat) <- paste0(exp, "_", used_channels)
  
  stopifnot(ncol(mat) == length(used_channels))
  
  # ---------------------
  # rowData
  # ---------------------
  row_data_exp <- DataFrame(
    Sequence        = exp_data$Sequence,
    ModifiedSeq     = exp_data$`Modified sequence`,
    Proteins        = exp_data$Proteins,
    LeadingProtein  = exp_data$`Leading proteins`,
    GeneNames       = exp_data$`Gene names`,
    Charge          = exp_data$Charge,
    PEP             = exp_data$PEP,
    Score           = exp_data$Score,
    RawFile         = exp_data$`Raw file`,
    Experiment      = exp_data$Experiment,
    Reverse         = exp_data$Reverse,
    Contaminant     = exp_data$`Potential contaminant`
  )
  
  rownames(row_data_exp) <- exp_data$id
  
  # ---------------------
  # colData
  # ---------------------
  col_data_exp <- sample_annotation %>%
    filter(Experiment == exp & TMT_channel %in% used_channels) %>%
    mutate(colname = paste0(Experiment, "_", TMT_channel)) %>%
    arrange(match(TMT_channel, used_channels)) %>%
    column_to_rownames("colname") %>%
    DataFrame()
  
  stopifnot(ncol(mat) == nrow(col_data_exp))
  
  # ---------------------
  # Create SummarizedExperiment
  # ---------------------
  se_exp <- SummarizedExperiment(
    assays  = list(intensity = mat),
    rowData = row_data_exp,
    colData = col_data_exp
  )
  
  qf[[exp]] <- se_exp
}

# ---------------------
# Inspect result
# ---------------------
qf
names(qf)
```

Inspect
```{r}
qf
summary(qf)
class(qf)
names(qf)
colData(qf)
rowData(qf)
anyNA(assay(qf))
```

## 4. Normalize psm intensities

```{r}
library(matrixStats)

psm_sets <- names(qf)

for (psm_name in psm_sets) {
  
  message("Processing normalization: ", psm_name)
  
  # -------------------------
  # 1) Extract raw assay
  # -------------------------
  raw_assay <- assay(qf[[psm_name]])
  raw_assay <- as.matrix(raw_assay)
  
  # -------------------------
  # 2) Fraction of column total
  # -------------------------
  frac_assay <- sweep(raw_assay, 2,
                      colSums(raw_assay, na.rm = TRUE),
                      FUN = "/")
  
  scaled_assay <- frac_assay * 1e6
  
  frac_name <- paste0(psm_name, "_frac")
  
  qf[[frac_name]] <- SummarizedExperiment(
    assays  = list(intensity = scaled_assay),
    rowData = rowData(qf[[psm_name]]),
    colData = colData(qf[[psm_name]])
  )
  
  qf <- addAssayLinkOneToOne(qf,
                             from = psm_name,
                             to   = frac_name)
  
  # -------------------------
  # 3) Add pseudocount
  # -------------------------
  frac_pc <- scaled_assay + 1e-6
  
  frac_pc_name <- paste0(psm_name, "_frac_pc")
  
  qf[[frac_pc_name]] <- SummarizedExperiment(
    assays  = list(intensity = frac_pc),
    rowData = rowData(qf[[frac_name]]),
    colData = colData(qf[[frac_name]])
  )
  
  qf <- addAssayLinkOneToOne(qf,
                             from = frac_name,
                             to   = frac_pc_name)
  
  # -------------------------
  # 4) log2 transform
  # -------------------------
  log_name <- paste0(psm_name, "_log")
  
  qf <- logTransform(
    qf,
    i    = frac_pc_name,
    name = log_name
  )
  
  # -------------------------
  # 5) Quantile normalize
  # -------------------------
  norm_name <- paste0(psm_name, "_norm")
  
  qf <- normalize(
    qf,
    i      = log_name,
    name   = norm_name,
    method = "quantiles"
  )
}

```

## 5. Aggregation
```{r}
norm_sets <- grep("_norm$", names(qf), value = TRUE)

for (norm_name in norm_sets) {
  
  message("Aggregating: ", norm_name)
  
  # -------------------------
  # PSM → Peptide
  # -------------------------
  pep_name <- paste0(norm_name, "_peptide")
  
  qf <- aggregateFeatures(
    object = qf,
    i      = norm_name,
    fcol   = "Sequence",
    name   = pep_name,
    fun    = function(x)
      matrixStats::colMedians(x, na.rm = TRUE)
  )
  
  # -------------------------
  # Peptide → Protein
  # -------------------------
  prot_name <- paste0(norm_name, "_protein")
  
  qf <- aggregateFeatures(
    object = qf,
    i      = pep_name,
    fcol   = "Proteins",
    name   = prot_name,
    fun    = function(x)
      matrixStats::colMedians(x, na.rm = TRUE)
  )
}
```

Inspect
```{r}
qf
summary(qf)
class(qf)
names(qf)
colData(qf)
rowData(qf)
anyNA(assay(qf))
library(dplyr)
library(purrr)

qfInfo <- tibble(
  assay  = names(qf),
  n_rows = map_int(qf, nrow),
  n_cols = map_int(qf, ncol)
) %>%
  arrange(assay)
```

## 11. Differential analysis

```{r, eval=FALSE, echo =FALSE, message=FALSE, warning=FALSE}
library(DEP)
library(dplyr)
library(tidyr)
library(S4Vectors)

protein_sets <- grep("_norm_protein$", names(qf), value = TRUE)

dep_results <- list()

for (prot_name in protein_sets) {
  
  message("Running DEP for: ", prot_name)
  
  protein_se <- qf[[prot_name]]
  
  # -------------------------
  # Build colData
  # -------------------------
  
  col_df <- as.data.frame(colData(protein_se))
  col_df$label     <- rownames(col_df)
  col_df$condition <- col_df$treatment
  col_df$replicate <- col_df$`repeat`
  
  # -------------------------
  # Build rowData (KEEP GeneNames)
  # -------------------------
  
  row_df <- as.data.frame(rowData(protein_se))
  
  # Make sure GeneNames column exists
  if (!"GeneNames" %in% colnames(row_df)) {
    stop("GeneNames column not found in rowData(protein_se)")
  }
  
  row_df$name  <- row_df$Proteins
  row_df$ID    <- paste0(row_df$Proteins, "_", seq_len(nrow(row_df)))
  row_df$gene  <- row_df$GeneNames   # 👈 IMPORTANT ADDITION
  
  assay_mat <- assay(protein_se)
  row_df <- cbind(row_df, assay_mat)
  
  columns <- (ncol(row_df) - ncol(assay_mat) + 1):ncol(row_df)
  
  # -------------------------
  # Make DEP object
  # -------------------------
  
  dep <- make_se(row_df, columns, col_df)
  
  # Impute
  dep <- impute(dep, method = "MinProb", q = 0.01)
  
  # Remove Pool
  dep <- dep[, dep$condition != "Pool"]
  
  # Control
  dep <- test_diff(
    dep,
    type    = "control",
    control = "C"
  )
  
  dep <- add_rejections(
    dep,
    alpha = 0.05,
    lfc   = 0
  )
  
  # Store results (NOW includes gene column)
  res <- get_results(dep)

# Extract gene names from DEP object
gene_df <- as.data.frame(rowData(dep))[, c("name", "gene")]

# Merge gene names back into results
res <- left_join(res, gene_df, by = "name")

dep_results[[prot_name]] <- res
}
```

```{r}
## Checking different thresholds
alpha_vals <- c(0.01, 0.05, 0.1)
lfc_vals   <- c(0, 0.5, 1)

sig_summary <- list()

for (prot_name in protein_sets) {

  message("Running DEP for: ", prot_name)
  
  protein_se <- qf[[prot_name]]
  
  col_df <- as.data.frame(colData(protein_se))
  col_df$label     <- rownames(col_df)
  col_df$condition <- col_df$treatment
  col_df$replicate <- col_df$`repeat`
  
  row_df <- as.data.frame(rowData(protein_se))
  row_df$name  <- row_df$Proteins
  row_df$ID    <- paste0(row_df$Proteins, "_", seq_len(nrow(row_df)))
  row_df$gene  <- row_df$GeneNames
  
  assay_mat <- assay(protein_se)
  row_df <- cbind(row_df, assay_mat)
  columns <- (ncol(row_df) - ncol(assay_mat) + 1):ncol(row_df)
  
  dep <- make_se(row_df, columns, col_df)
  dep <- impute(dep, method ="MinProb", q=0.01)
  dep <- dep[, dep$condition != "Pool"]
  
  dep <- test_diff(dep, type="control", control="C")
  
  # Test multiple alpha / lfc combinations
  for(a in alpha_vals){
    for(l in lfc_vals){
      
      dep_tmp <- add_rejections(dep, alpha=a, lfc=l)
      
      res <- get_results(dep_tmp)
      
      n_sig <- sum(res$significant, na.rm=TRUE)
      
      sig_summary[[length(sig_summary)+1]] <- data.frame(
        dataset = prot_name,
        alpha   = a,
        lfc     = l,
        significant_proteins = n_sig
      )
    }
  }
}

sig_summary <- bind_rows(sig_summary)
sig_summary
ggplot(sig_summary, aes(x=lfc, y=significant_proteins, color=factor(alpha))) +
  geom_line() +
  geom_point(size=3) +
  facet_wrap(~dataset) +
  labs(
    color="alpha",
    x="log2FC threshold",
    y="Number of significant proteins"
  ) +
  theme_bw()
```


## 12. Upset plot 1


```{r}
library(dplyr)
library(UpSetR)
library(viridisLite)

vcols <- viridis(6)

# Google colors for sets
treatment_colors <- c("A"=vcols[2],"V"=vcols[4],"VA"=vcols[6])

make_upset <- function(res, prefix) {

  upset_df <- res %>%
    dplyr::select(name, A_vs_C_significant, V_vs_C_significant, VA_vs_C_significant) %>%
    dplyr::mutate(
      A = as.integer(A_vs_C_significant),
      V = as.integer(V_vs_C_significant),
      VA = as.integer(VA_vs_C_significant)
    ) %>%
    dplyr::select(name, A, V, VA)

  upset_df <- upset_df[rowSums(upset_df[, -1]) > 0, ]

  if(nrow(upset_df) == 0) {
    message("No significant proteins to plot for ", prefix)
    return(NULL)
  }

  set_order <- c("V","A","VA")

  UpSetR::upset(
    upset_df,
     # Set structure
    sets = set_order,
    nsets = 3,
    nintersects = NA,
    order.by = "freq",
    decreasing = TRUE,
    main.bar.color = vcols[5],
    sets.bar.color = treatment_colors[set_order],  # match order
    # Matrix appearance
    matrix.color = vcols[3],
    point.size = 4,
    line.size = 1.2,
    mb.ratio = c(0.65, 0.35),
    # Text scaling
    text.scale = c(1.4, 1.6, 1, 1.2, 1.6, 1.4),
    # Intersection numbers
    show.numbers = "yes",
    number.angles = 30,
    # Background shading
    shade.color = vcols[1],
    shade.alpha = 0.2,
    mainbar.y.label = "Number of proteins",
    sets.x.label = "Proteins per treatment",
    keep.order = TRUE
  )
}
    
# kit_titles <- c(
#   "Kit4_2_norm_protein" = "Short-term living cell CoPISA",
#   "Kit4_3_norm_protein" = "Long-term living cell CoPISA",
#   "Kit4_4_norm_protein" = "Short-term Cell lysate CoPISA",
#   "Kit4_5_norm_protein" = "Long-term expression proteomics"
# )


up1 <- make_upset(dep_results[[1]], "Kit4_2_norm_protein")
up2 <- make_upset(dep_results[[2]], "Kit4_3_norm_protein")
up3 <- make_upset(dep_results[[3]], "Kit4_4_norm_protein")
up4 <- make_upset(dep_results[[4]], "Kit4_5_norm_protein")
up2
up3
up1
up4


# Save up1 as PNG
png("up1.png", width = 10, height = 10, units = "cm", res = 300)
print(up1)  # must use print() for ComplexUpset
dev.off()

# Repeat for other plots
png("up2.png", width = 10, height = 10, units = "cm", res = 300)
print(up2)
dev.off()

png("up3.png", width = 10, height = 10, units = "cm", res = 300)
print(up3)
dev.off()

png("up4.png", width = 10, height = 10, units = "cm", res = 300)
print(up4)
dev.off()
```


## 13. Upset plot 2

```{r}
library(dplyr)
library(tidyr)
library(UpSetR)
library(viridis)
library(purrr)  # for reduce()

# ---- 1. Prepare combined VA significant dataframe ----
upset_combined <- lapply(names(dep_results), function(nm) {
  res <- dep_results[[nm]]
  
  res %>%
    select(name = gene, VA_vs_C_significant) %>%
    mutate(!!nm := as.integer(VA_vs_C_significant)) %>%
    select(-VA_vs_C_significant)
}) %>%
  purrr::reduce(full_join, by = "name")  # combine all 4 datasets by protein name

# Replace NAs with 0 (not significant)
upset_combined[is.na(upset_combined)] <- 0

# ---- 2. Map kit names to descriptive titles ----
kit_titles <- c(
  "Kit4_2_norm_protein" = "Short-term living cell CoPISA",
  "Kit4_3_norm_protein" = "Long-term living cell CoPISA",
  "Kit4_4_norm_protein" = "Short-term Cell lysate CoPISA",
  "Kit4_5_norm_protein" = "Long-term expression proteomics"
)

# Replace column names with descriptive names
colnames(upset_combined) <- c("name", kit_titles[colnames(upset_combined)[-1]])

# ---- 3. Define set order and colors ----
set_order <- kit_titles  # use descriptive names

pcols <- plasma(6)
treatment_colors <- c(
  "Short-term living cell CoPISA"    = pcols[2],
  "Long-term living cell CoPISA"     = pcols[3],
  "Short-term Cell lysate CoPISA"    = pcols[4],
  "Long-term expression proteomics"  = pcols[5]
)

# ---- 4. Create UpSet plot ----

upset_VA <- UpSetR::upset(
  upset_combined,
  
  sets = set_order,
  nsets = length(set_order),
  nintersects = NA,
  order.by = "freq",
  decreasing = TRUE,
  
  main.bar.color = vcols[6],                     # intersection bars
  sets.bar.color = treatment_colors[set_order],  # treatment bars
  matrix.color = vcols[3],                       # intersection points
  point.size = 4,
  line.size = 1.2,
  mb.ratio = c(0.65, 0.35),
  
  text.scale = c(1.4, 1.6, 1, 1.2, 1.6, 1.4),
  
  show.numbers = "yes",
  number.angles = 30,
  
  shade.color = vcols[1],
  shade.alpha = 0.2,
  
  mainbar.y.label = "Number of proteins",
  sets.x.label = "Proteins per VA \ntreatment in all four samples",
  keep.order = TRUE
)

```

```{r}
# ---- 1. Prepare combined V significant dataframe ----
upset_combined <- lapply(names(dep_results), function(nm) {
  res <- dep_results[[nm]]
  
  res %>%
    select(name = gene, V_vs_C_significant) %>%
    mutate(!!nm := as.integer(V_vs_C_significant)) %>%
    select(-V_vs_C_significant)
}) %>%
  purrr::reduce(full_join, by = "name")  # combine all 4 datasets by protein name

# Replace NAs with 0 (not significant)
upset_combined[is.na(upset_combined)] <- 0

# ---- 2. Map kit names to descriptive titles ----
kit_titles <- c(
  "Kit4_2_norm_protein" = "Short-term living cell CoPISA",
  "Kit4_3_norm_protein" = "Long-term living cell CoPISA",
  "Kit4_4_norm_protein" = "Short-term Cell lysate CoPISA",
  "Kit4_5_norm_protein" = "Long-term expression proteomics"
)

# Replace column names with descriptive names
colnames(upset_combined) <- c("name", kit_titles[colnames(upset_combined)[-1]])

# ---- 3. Define set order and colors ----
set_order <- kit_titles  # use descriptive names

pcols <- plasma(6)
treatment_colors <- c(
  "Short-term living cell CoPISA"    = pcols[2],
  "Long-term living cell CoPISA"     = pcols[3],
  "Short-term Cell lysate CoPISA"    = pcols[4],
  "Long-term expression proteomics"  = pcols[5]
)

# ---- 4. Create UpSet plot ----

upset_V <- UpSetR::upset(
  upset_combined,
  
  sets = set_order,
  nsets = length(set_order),
  nintersects = NA,
  order.by = "freq",
  decreasing = TRUE,
  
  main.bar.color = vcols[4],                     # intersection bars
  sets.bar.color = treatment_colors[set_order],  # treatment bars
  matrix.color = vcols[3],                       # intersection points
  point.size = 4,
  line.size = 1.2,
  mb.ratio = c(0.65, 0.35),
  
  text.scale = c(1.4, 1.6, 1, 1.2, 1.6, 1.4),
  
  show.numbers = "yes",
  number.angles = 30,
  
  shade.color = vcols[1],
  shade.alpha = 0.2,
  
  mainbar.y.label = "Number of proteins",
  sets.x.label = "Proteins per V \ntreatment in all four samples",
  keep.order = TRUE
)

```

```{r}
# ---- 1. Prepare combined A significant dataframe ----
upset_combined <- lapply(names(dep_results), function(nm) {
  res <- dep_results[[nm]]
  
  res %>%
    select(name = gene, A_vs_C_significant) %>%
    mutate(!!nm := as.integer(A_vs_C_significant)) %>%
    select(-A_vs_C_significant)
}) %>%
  purrr::reduce(full_join, by = "name")  # combine all 4 datasets by protein name

# Replace NAs with 0 (not significant)
upset_combined[is.na(upset_combined)] <- 0

# ---- 2. Map kit names to descriptive titles ----
kit_titles <- c(
  "Kit4_2_norm_protein" = "Short-term living cell CoPISA",
  "Kit4_3_norm_protein" = "Long-term living cell CoPISA",
  "Kit4_4_norm_protein" = "Short-term Cell lysate CoPISA",
  "Kit4_5_norm_protein" = "Long-term expression proteomics"
)

# Replace column names with descriptive names
colnames(upset_combined) <- c("name", kit_titles[colnames(upset_combined)[-1]])

# ---- 3. Define set order and colors ----
set_order <- kit_titles  # use descriptive names

pcols <- plasma(6)
treatment_colors <- c(
  "Short-term living cell CoPISA"    = pcols[2],
  "Long-term living cell CoPISA"     = pcols[3],
  "Short-term Cell lysate CoPISA"    = pcols[4],
  "Long-term expression proteomics"  = pcols[5]
)

# ---- 4. Create UpSet plot ----

upset_A <- UpSetR::upset(
  upset_combined,
  
  sets = set_order,
  nsets = length(set_order),
  nintersects = NA,
  order.by = "freq",
  decreasing = TRUE,
  
  main.bar.color = vcols[2],                     # intersection bars
  sets.bar.color = treatment_colors[set_order],  # treatment bars
  matrix.color = vcols[3],                       # intersection points
  point.size = 4,
  line.size = 1.2,
  mb.ratio = c(0.65, 0.35),
  
  text.scale = c(1.4, 1.6, 1, 1.2, 1.6, 1.4),
  
  show.numbers = "yes",
  number.angles = 30,
  
  shade.color = vcols[1],
  shade.alpha = 0.2,
  
  mainbar.y.label = "Number of proteins",
  sets.x.label = "Proteins per A \ntreatment in all four samples",
  keep.order = TRUE
)

```


```{r}
png("upset_VA.png", width = 15, height = 10, units = "cm", res = 300)
upset_VA
dev.off()

png("upset_V.png", width = 15, height = 10, units = "cm", res = 300)
upset_V
dev.off()

png("upset_A.png", width = 15, height = 10, units = "cm", res = 300)
upset_A
dev.off()

upset_VA
upset_V
upset_A
```
### Extract gene name of common sets
```{r}
# library(dplyr)
# 
# set_cols <- colnames(upset_combined)[-1]
# n_sets <- length(set_cols)

# Create combinations
comb_df <- upset_combined %>%
  mutate(
    combination = apply(select(., all_of(set_cols)), 1, function(x) {
      paste(set_cols[x == 1], collapse = " & ")
    }),
    n_members = rowSums(select(., all_of(set_cols)))
  ) %>%
  filter(combination != "")


#---- OPTION 2 (alternative): only shared sets (>=2) ----
(intersection_labels <- comb_df %>%
  filter(n_members >= 2) %>%
  group_by(combination) %>%
  summarise(genes = paste(name, collapse = ", "), .groups = "drop")
)

```



## 14. Volcano plot
```{r}
# library(dplyr)
# library(tidyr)
# library(ggplot2)
# library(viridis)
# library(ggrepel)

make_volcano <- function(res, prefix) {
  
  pval_thresh <- 0.05
  lfc_thresh  <- 1
  
  # Kit name mapping
  kit_labels <- c(
    "Kit4_2_norm_protein" = "Intact_short_term",
    "Kit4_3_norm_protein" = "Intact_long_term",
    "Kit4_4_norm_protein" = "Extract_short_term",
    "Kit4_5_norm_protein" = "Expression_long_term"
  )
  
  plot_name <- kit_labels[prefix]
  if (is.na(plot_name)) plot_name <- prefix
  
  # Prepare long format
  long_df <- res %>%
    select(name, gene,
           A_vs_C_ratio, A_vs_C_p.val,
           V_vs_C_ratio, V_vs_C_p.val,
           VA_vs_C_ratio, VA_vs_C_p.val) %>%
    pivot_longer(
      cols = -c(name, gene),
      names_to = c("Treatment", ".value"),
      names_pattern = "(.*)_vs_C_(ratio|p.val)"
    )  %>%
    mutate(log2FC = ratio,
           pval   = p.val) %>%
    mutate(
      Treatment = recode(Treatment,
                         "A"  = "Azacytidine",
                         "VA" = "Azacytidine + Venetoclax",
                         "V"  = "Venetoclax"),
      negLogP = -log10(pval),
      Significant = (pval < pval_thresh & abs(log2FC) > lfc_thresh),
      gene = ifelse(is.na(gene) | gene == "", name, gene),
      gene = sub(";.*", "", gene)
    )
  
  # Determine shared vs unique (ONLY among significant proteins)
  shared_info <- long_df %>%
    filter(Significant) %>%
    group_by(name) %>%
    summarise(n_sig = n(), .groups = "drop")
  
  long_df <- long_df %>%
    left_join(shared_info, by = "name") %>%
    mutate(
      Category = case_when(
        Significant & n_sig > 1  ~ "Shared significant protein",
        Significant & n_sig == 1 ~ "Unique significant protein",
        TRUE ~ "Not significant"
      )
    )
  
  # 🔥 Select top 5 significant proteins per treatment (smallest p-value)
  label_df <- long_df %>%
    filter(Significant) %>%
    group_by(Treatment) %>%
    arrange(pval, .by_group = TRUE) %>%
    slice_head(n = 5) %>%
    ungroup()
  
  # 🔥 MODIFIED: define custom viridis colors
  library(viridis)
  vcols <- viridis(6)
  treatment_colors <- c(
    "Azacytidine"             = vcols[2],
    "Venetoclax"              = vcols[4],
    "Azacytidine + Venetoclax"= vcols[6]
  )
  
  p <- ggplot(long_df, aes(x = log2FC, y = negLogP)) +
    geom_point(aes(color = Treatment,
                   fill  = Category,
                   alpha = Significant),
               shape = 21,
               size = 1.2,
               stroke = 1
               ) +
  
    scale_alpha_manual(values = c(
      "TRUE"  = 1,
      "FALSE" = 0.15
    ), guide = "none") +
    
    geom_text_repel(
                data = label_df,
                aes(label = gene, color = Treatment),
                size = 3,
                box.padding = 0.3,
                max.overlaps = 20,
                force = 1,
                segment.size = 0.3,
                segment.alpha = 0.5
    ) +
    
    geom_vline(xintercept = c(-lfc_thresh, lfc_thresh),
               linetype = "dashed") +
    geom_hline(yintercept = -log10(pval_thresh),
               linetype = "dashed") +
    
    # 🔥 MODIFIED: use manual color assignment instead of viridis_d()
    scale_color_manual(values = treatment_colors) +
    
    scale_fill_manual(values = c(
      "Shared significant protein" = "black",
      "Unique significant protein" = "gray60",
      "Not significant" = "white"
    )) +
    
    theme_minimal(base_size = 13) +
    labs(
      title = paste("Volcano Plot -", plot_name),
      x = "log2 Fold Change",
      y = "-log10(p-value)"
    ) +
    theme(
      plot.title = element_text(face = "bold"),
      legend.position = "bottom"
    )
  
  return(p)
}


```



```{r}
#library(patchwork)

# ---- 1. Define consistent axis limits ----
all_long <- lapply(names(dep_results), function(nm) {
  res <- dep_results[[nm]]
  
  res %>%
    select(A_vs_C_ratio, A_vs_C_p.val,
           V_vs_C_ratio, V_vs_C_p.val,
           VA_vs_C_ratio, VA_vs_C_p.val) %>%
    pivot_longer(everything(),
                 names_to = c("Treatment", ".value"),
                 names_pattern = "(.*)_vs_C_(.*)") %>%
    dplyr::rename(log2FC = ratio,
           pval   = p.val) %>%
    mutate(negLogP = -log10(pval))
})

all_long <- bind_rows(all_long)

x_limits <- range(all_long$log2FC, na.rm = TRUE)
y_limits <- range(all_long$negLogP, na.rm = TRUE)

# ---- 2. Custom panel titles ----
custom_titles <- c(
  "Kit4_2_norm_protein" = "Short-term living cell CoPISA",
  "Kit4_3_norm_protein" = "Long-term living cell CoPISA",
  "Kit4_4_norm_protein" = "Short-term Cell lysate CoPISA",
  "Kit4_5_norm_protein" = "Long-term expression protoemics"
)

# ---- 3. Create plots with modifications ----
plots <- lapply(names(dep_results), function(nm) {
  
  p <- make_volcano(dep_results[[nm]], nm) +
  coord_cartesian(xlim = c(-4, 4), ylim = y_limits) +
  labs(title = custom_titles[nm]) +
  theme(
    panel.border = element_rect(
      colour = "black",
      fill = NA,
      linewidth = 0.6
    )
  )
  
  return(p)
})

# ---- 4. Remove redundant axis titles ----
# Layout:
# [1] [2]
# [3] [4]

p1 <- plots[[3]] + theme(axis.title.x = element_blank())
p2 <- plots[[4]] + theme(axis.title.x = element_blank(),
                         axis.title.y = element_blank())
p3 <- plots[[1]] 
p4 <- plots[[2]] + theme(axis.title.y = element_blank())

# ---- 5. Combine with shared legend ----
library(patchwork)

combined_plot <- (
  (p1 + p2) /
  (p3 + p4)
) +
  plot_layout(guides = "collect") &
  theme(
    plot.title = element_text(face = "bold"),
    legend.position = "bottom",
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 9),
    legend.key.size = unit(0.6, "lines")
  )

combined_plot
ggsave("Volcano_4panel_combined.png",
       combined_plot,
       width = 23.5,
        height = 14.85,
        units = "cm",
        dpi = 300)
```


## 15. Gene of interest
```{r}
library(dplyr)

# Create a summary of BCL2 across all kits
bcl2_summary <- lapply(names(dep_results), function(nm) {
  res <- dep_results[[nm]]
  
  # Make sure gene column exists
  gene_col <- if ("gene" %in% colnames(res)) "gene" else "name"
  
  res %>%
    filter(grepl("^map7$", !!sym(gene_col), ignore.case = TRUE)) %>%
    mutate(Kit = nm)
}) %>% bind_rows()

# View summary
bcl2_summary


# UniProt ID to check
uniprot_id <- "P10415"

# Search across all dep_results
uniprot_summary <- lapply(names(dep_results), function(nm) {
  res <- dep_results[[nm]]
  
  res %>%
    filter(grepl(uniprot_id, name, ignore.case = TRUE)) %>%
    mutate(Kit = nm)
}) %>% bind_rows()

# View result
uniprot_summary

uniprot_id <- "Q24K09"

# Loop through all protein-level assays
for (nm in grep("_norm_protein$", names(qf), value = TRUE)) {
  protein_se <- qf[[nm]]
  
  found <- rowData(protein_se)$Proteins %>% 
           grepl(uniprot_id, ., ignore.case = TRUE)
  
  cat("Kit:", nm, "- P10415 present?", any(found), "\n")
  
  if (any(found)) {
    # Show the rows with this UniProt ID
    print(as.data.frame(rowData(protein_se)[found, ]))
  }
}
```

## 16. FC2FC common protein for VA
```{r}
kit_order <- c(
  "Kit4_2_norm_protein",  # Living cell short-term
  "Kit4_3_norm_protein",  # Living cell long-term
  "Kit4_4_norm_protein",  # Cell lysate short-term
  "Kit4_5_norm_protein"   # Expression long-term
)
```


```{r}
library(dplyr)

# Collect all VA log2FC values across all kits
all_fc <- lapply(kit_order, function(nm) {
  dep_results[[nm]]$VA_vs_C_ratio
}) %>% unlist()

lim <- max(abs(all_fc), na.rm = TRUE)
axis_limits <- c(-4, 4)
```

```{r}
# library(dplyr)
# library(ggplot2)
# library(viridis)

make_fc2fc_pair <- function(res1, res2, name1, name2,
                            expression_res = NULL) {
  
  df1 <- res1 %>%
    select(name,
           VA_vs_C_ratio,
           VA_vs_C_p.val) %>%
    rename(log2FC_1 = VA_vs_C_ratio,
           pval_1   = VA_vs_C_p.val)
  
  df2 <- res2 %>%
    select(name,
           VA_vs_C_ratio,
           VA_vs_C_p.val) %>%
    rename(log2FC_2 = VA_vs_C_ratio,
           pval_2   = VA_vs_C_p.val)
  
  merged <- inner_join(df1, df2, by = "name") %>%
    mutate(
      mean_negLogP = -log10((pval_1 + pval_2) / 2)
    )
  
  # Regression
  fit <- lm(log2FC_2 ~ log2FC_1, data = merged)
  slope <- coef(fit)[2]
  r2    <- summary(fit)$r.squared
  pval_reg <- summary(fit)$coefficients[2,4]
  
  label_text <- paste0(
    "Slope = ", round(slope, 3),
    "\nR² = ", round(r2, 3),
    "\nP = ", signif(pval_reg, 3)
  )
  
  p <- ggplot(merged, aes(log2FC_1, log2FC_2)) +
    
    geom_point(aes(fill = mean_negLogP),
               shape = 21,
               size = 1.6,
               alpha = 0.6,
               color = "gray40",  
               stroke = 0.6) +
    
    geom_smooth(method = "lm",
                se = FALSE,
                color = "red",   
                linewidth = 1.2) +
    
    annotate("text",
             x = axis_limits[1] * 0.9,
             y = axis_limits[2] * 0.9,
             label = label_text,
             hjust = 0,
             vjust = 1,
             size = 3,
             color = "blue") +
    
    scale_fill_viridis_c(
      option = "D",
      direction = -1,
      name = "-log10(p-value)"
    ) +
    
    coord_fixed(
      xlim = axis_limits,
      ylim = axis_limits,
      ratio = 1
    ) +
    
    theme_minimal(base_size = 11) +
    
    labs(
      title = paste(custom_titles[name1], "vs", custom_titles[name2]),
      x = paste0("log2FC (", custom_titles[name1], ")"),
      y = paste0("log2FC (", custom_titles[name2], ")")
    ) +
    
    theme(
      plot.title = element_text(face = "bold", size = 7),
      legend.position = "bottom",
      legend.text = element_text(size = 5),
      legend.title = element_text(size = 6),
      axis.title = element_text(size = 6)
    )
  
  return(p)
}
```




```{r}
p4 <- make_fc2fc_pair(dep_results[[kit_order[1]]],
                      dep_results[[kit_order[4]]],
                      kit_order[1], kit_order[4])

p5 <- make_fc2fc_pair(dep_results[[kit_order[3]]],
                      dep_results[[kit_order[4]]],
                      kit_order[3], kit_order[4])

p6 <- make_fc2fc_pair(dep_results[[kit_order[2]]],
                      dep_results[[kit_order[4]]],
                      kit_order[2], kit_order[4])

triple_plot_2 <- (p4 | p5 | p6) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

triple_plot_2

ggsave("FC2FC2Expression.png",
       triple_plot_2,
       width = 21,
        height = 10,
        units = "cm",
        dpi = 300)

```


```{r}
# library(ggplot2)
# library(dplyr)
# library(viridis)
# library(RColorBrewer)

make_fc2fc_pair1 <- function(res1, res2, res3, name1, name2, name3, axis_limits = c(-3, 3)) {
  
  # Merge the VA log2FC
  df1 <- res1 %>%
    select(name, VA_vs_C_ratio, gene) %>%
    rename(log2FC_1 = VA_vs_C_ratio)
  
  df2 <- res2 %>%
    select(name, VA_vs_C_ratio, gene) %>%
    rename(log2FC_2 = VA_vs_C_ratio)
  
  df3 <- res3 %>%
    select(name, VA_vs_C_ratio, gene) %>%
    rename(log2FC_3 = VA_vs_C_ratio)
  
  # Merge all three by 'name'
  merged <- df1 %>%
    full_join(df2, by = "name") %>%
    full_join(df3, by = "name")
  
  # Before labeling, create a gene column
merged <- merged %>%
  separate(gene, into = "gene", sep = ";", extra = "drop", fill = "right")
# Subset for labeling genes with large FC
label_df <<- merged %>%
  filter(abs(log2FC_1) > 1.5 | abs(log2FC_2) > 1.5)
  
  # Plot
  p <- ggplot(merged, aes(x = log2FC_1, y = log2FC_2)) +
    geom_point(aes(fill = log2FC_3),
               shape = 21, size = 1.5, alpha = 0.8, stroke = 0.5, color = "gray80") +
    
    # Add gene labels for large FC points
    ggrepel::geom_text_repel(
                    data = label_df,
                    aes(label = gene),   # use gene name instead of UniProt ID
                    size = 2,
                    box.padding = 0.3,
                    max.overlaps = 20,   # max number of labels allowed
                    force = 1, 
                    segment.color = "gray50"
                  ) +
    
    # Diverging color scale for FC of third kit
    scale_fill_gradient2(
      low = "blue",
      mid = "white",
      high = "red",
      midpoint = 0,
      limits = c(-0.3, 0.3),
      name = paste0("log2FC (", name3, ")"),
      oob = scales::squish
    ) +
    
    coord_fixed(xlim = axis_limits, ylim = axis_limits) +
    theme_minimal(base_size = 11) +
    labs(
      title = paste(name1, "vs", name2),
      x = paste0("log2FC (", name1, ")"),
      y = paste0("log2FC (", name2, ")")
    ) +
    theme(
      plot.title = element_text(face = "bold", size = 7),
      legend.position = "bottom",
      legend.text = element_text(size = 5),
      legend.title = element_text(size = 6),
      axis.title = element_text(size = 6)
    ) +
    guides(
      fill = guide_colorbar(barwidth = 3)
    )
  
  return(p)
}

```

```{r}
# First three plots (FC color)
p1 <- make_fc2fc_pair1(
  dep_results[["Kit4_3_norm_protein"]],
  dep_results[["Kit4_2_norm_protein"]],
  dep_results[["Kit4_5_norm_protein"]],
  "Long-term living cell CoPISA",
  "Short-term living cell CoPISA",
  "Long-term expression protoemics"
)


p2 <- make_fc2fc_pair1(
  dep_results[["Kit4_3_norm_protein"]],
  dep_results[["Kit4_4_norm_protein"]],
  dep_results[["Kit4_5_norm_protein"]],
  "Long-term living cell CoPISA",
  "Short-term Cell lysate CoPISA",
  "Long-term expression protoemics"
)


triple_plot_1 <- (p1 | p3) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

triple_plot_1

ggsave("FC2FCLongtoShort.png",
       triple_plot_1,
       width = 15,
        height = 10,
        units = "cm",
        dpi = 300)

```

## 17. Merge Four Proteomics Analyses by Fisher's Method for Combining P-values

```{r}
# -----------------------------
# 1. Prepare merged dataset
# -----------------------------

# Get common proteins across all datasets
common_proteins <- Reduce(intersect, list(
  dep_results$Kit4_2_norm_protein$gene,
  dep_results$Kit4_3_norm_protein$gene,
  dep_results$Kit4_4_norm_protein$gene,
  dep_results$Kit4_5_norm_protein$gene
))

# Filter datasets for common proteins
dep1 <- dep_results$Kit4_3_norm_protein %>% filter(gene %in% common_proteins) # "Long-term living cell CoPISA",
dep2 <- dep_results$Kit4_2_norm_protein %>% filter(gene %in% common_proteins) # "Short-term living cell CoPISA"
dep3 <- dep_results$Kit4_4_norm_protein %>% filter(gene %in% common_proteins) # "Short-term Cell lysate CoPISA"
dep4 <- dep_results$Kit4_5_norm_protein %>% filter(gene %in% common_proteins) # "Long-term expression proteomics"

deps <- list(dep1, dep2, dep3, dep4)

# Merge datasets by gene name and suffix columns with dataset index
merged <- imap(deps, function(df, i) {
  df %>%
    select(name = gene, ends_with("_p.val"), ends_with("_ratio")) %>%
    rename_with(~ paste0(.x, ".", i), -name)  # i is from imap
}) %>%
  purrr::reduce(inner_join, by = "name")

# -----------------------------
# 2. Helper functions
# -----------------------------

# Fisher method for combining p-values
combine_fisher <- function(...) {
  pvals <- c(...)
  pvals[is.na(pvals) | pvals == 0] <- 1
  chisq <- -2 * sum(log(pvals))
  df <- 2 * length(pvals)
  pchisq(chisq, df = df, lower.tail = FALSE)
}

# Harmonic mean for absolute fold change (pairwise)
hm2 <- function(x, y) {
  2 / (1/abs(x) + 1/abs(y))
}

# -----------------------------
# 3. Compute pairwise Fisher and HM fold changes
# -----------------------------

merged <- merged %>%
  rowwise() %>%
  mutate(
    # ---- A vs C ----
    Fisher_A_12 = combine_fisher(A_vs_C_p.val.1, A_vs_C_p.val.2),
    Fisher_A_13 = combine_fisher(A_vs_C_p.val.1, A_vs_C_p.val.3),
    Fisher_A_23 = combine_fisher(A_vs_C_p.val.2, A_vs_C_p.val.3),

    hmAbsFC_A_12 = hm2(A_vs_C_ratio.1, A_vs_C_ratio.2),
    hmAbsFC_A_13 = hm2(A_vs_C_ratio.1, A_vs_C_ratio.3),
    hmAbsFC_A_23 = hm2(A_vs_C_ratio.2, A_vs_C_ratio.3),

    # ---- V vs C ----
    Fisher_V_12 = combine_fisher(V_vs_C_p.val.1, V_vs_C_p.val.2),
    Fisher_V_13 = combine_fisher(V_vs_C_p.val.1, V_vs_C_p.val.3),
    Fisher_V_23 = combine_fisher(V_vs_C_p.val.2, V_vs_C_p.val.3),

    hmAbsFC_V_12 = hm2(V_vs_C_ratio.1, V_vs_C_ratio.2),
    hmAbsFC_V_13 = hm2(V_vs_C_ratio.1, V_vs_C_ratio.3),
    hmAbsFC_V_23 = hm2(V_vs_C_ratio.2, V_vs_C_ratio.3),

    # ---- VA vs C ----
    Fisher_VA_12 = combine_fisher(VA_vs_C_p.val.1, VA_vs_C_p.val.2),
    Fisher_VA_13 = combine_fisher(VA_vs_C_p.val.1, VA_vs_C_p.val.3),
    Fisher_VA_23 = combine_fisher(VA_vs_C_p.val.2, VA_vs_C_p.val.3),

    hmAbsFC_VA_12 = hm2(VA_vs_C_ratio.1, VA_vs_C_ratio.2),
    hmAbsFC_VA_13 = hm2(VA_vs_C_ratio.1, VA_vs_C_ratio.3),
    hmAbsFC_VA_23 = hm2(VA_vs_C_ratio.2, VA_vs_C_ratio.3)
  ) %>%
  ungroup()

# -----------------------------
# 4. Geometric mean cutoff for |log2FC|
# -----------------------------
# cutoffs_log2 <- c(0.3, 0.3)
# cutoffs_linear <- 2^cutoffs_log2
# gm_cutoff_linear <- exp(mean(log(cutoffs_linear)))
# gm_cutoff_log2 <- log2(gm_cutoff_linear)
gm_cutoff_log2 <- quantile(abs(plot_data$gmAbsFC), 0.75, na.rm = TRUE)
# -----------------------------
# 5. Prepare data for plotting
# -----------------------------
plot_data <- merged %>%
  mutate(
    # Convert HM |log2FC| to linear
    gmAbsFC_VA_12 = 2^hmAbsFC_VA_12,
    gmAbsFC_VA_13 = 2^hmAbsFC_VA_13,
    gmAbsFC_VA_23 = 2^hmAbsFC_VA_23
  ) %>%
  rowwise() %>%
  mutate(
    # Geometric mean in linear space (safeguard)
    gmAbsFC_VA_12 = exp(mean(log(c(gmAbsFC_VA_12)), na.rm = TRUE)),
    gmAbsFC_VA_13 = exp(mean(log(c(gmAbsFC_VA_13)), na.rm = TRUE)),
    gmAbsFC_VA_23 = exp(mean(log(c(gmAbsFC_VA_23)), na.rm = TRUE))
  ) %>%
  ungroup() %>%
  mutate(
    # Back to log2 for plotting
    gmAbsFC_VA_12 = log2(gmAbsFC_VA_12),
    gmAbsFC_VA_13 = log2(gmAbsFC_VA_13),
    gmAbsFC_VA_23 = log2(gmAbsFC_VA_23)
  ) %>%
  select(
    name,
    gmAbsFC_VA_12, Fisher_VA_12,
    gmAbsFC_VA_13, Fisher_VA_13,
    gmAbsFC_VA_23, Fisher_VA_23
  ) %>%
  pivot_longer(
    cols = -name,
    names_to = c(".value", "Condition"),
    names_pattern = "(gmAbsFC|Fisher)_(VA_\\d+)"
  ) %>%
  mutate(
    Condition = recode(Condition,
      "VA_12" = "Long-term living cell CoPISA vs \nShort-term living cell CoPISA",
      "VA_13" = "Long-term living cell CoPISA vs \nShort-term Cell lysate CoPISA",
      "VA_23" = "Short-term living cell CoPISA vs \nShort-term Cell lysate CoPISA"
    ),
    significant = ifelse(Fisher < 0.05 & gmAbsFC > gm_cutoff_log2,
                         "Significant", "Not significant")
  )

# -----------------------------
# 6. Volcano-style plot
# -----------------------------
library(ggplot2)
library(ggrepel)
# --- 1. Remove multiple gene names and keep the first one to show 
plot_data <- plot_data %>%
  mutate(name_plot = sub(";.*$", "", name))  # take everything before the first semicolon

# --- 2. Add dep4 fold-change to plot_data ---
dep4_fc <- dep_results$Kit4_5_norm_protein %>%
  filter(gene %in% common_proteins) %>%
  select(name = gene, FC_dep4 = VA_vs_C_ratio)  # or whichever comparison you want

plot_data <- plot_data %>%
  left_join(dep4_fc, by = c("name" = "name")) %>%
  mutate(
    # keep only first gene if multiple genes
    name_plot = sub(";.*$", "", name)
  )

# Volcano-style plot with cutoff lines
spray <- ggplot(plot_data, aes(x = gmAbsFC, y = -log10(Fisher))) +
  geom_point(aes(
    color = ifelse(abs(gmAbsFC) > gm_cutoff_log2 & Fisher < 0.05,
                   dep_results$Kit4_5_norm_protein$A_vs_C_ratio[match(name, dep_results$Kit4_5_norm_protein$gene)],
                   NA)
  ), alpha = 0.8, size = 2) +
  scale_color_gradient2(
    low = "blue", mid = "gray90", high = "red",
    midpoint = 0,
    na.value = "gray70",
    name = "Long-term expression proteomics log2FC"
  ) +
  geom_text_repel(
    data = subset(plot_data, abs(gmAbsFC) > gm_cutoff_log2 & Fisher < 0.05),
    aes(label = sub(";.*$", "", name)),
    size = 3,
    max.overlaps = 20,
    box.padding = 0.4,
    point.padding = 0.3,
    segment.color = "grey50",
    show.legend = FALSE
  ) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "blue") +
  geom_vline(xintercept = gm_cutoff_log2, linetype = "dashed", color = "blue") +
  facet_wrap(~Condition, scales = "free_y") +
  theme_bw(base_size = 13) +
  theme(
    legend.position = "top",
    strip.background = element_rect(fill = "white", color = "black"),
    panel.grid = element_blank()
  ) +
  labs(x = "Geometric Mean |log2 Fold Change|",
       y = "-log10(Fisher Combined p-value)"
  )

spray
# Save your plot 'spray' as PNG
# ggsave(
#   filename = "sprayPlot.png",
#   plot = spray,          # your ggplot object
#   width = 17,            # width in cm
#   height = 9,            # height in cm
#   units = "cm",
#   dpi = 300
# )

```
# 18. R1-R4 group of proteins
```{r}
library(dplyr)

short_term_subset <- plot_data %>%
  filter(
    Condition == "Short-term living cell CoPISA vs \nShort-term Cell lysate CoPISA"
  )

short_term_subset <- short_term_subset %>%
  mutate(
    absent_in_long_term = !name %in% (
      plot_data %>%
        filter(
          Condition %in% c(
            "Long-term living cell CoPISA vs \nShort-term Cell lysate CoPISA",
            "Long-term living cell CoPISA vs \nShort-term living cell CoPISA"
          ),
          significant == "Significant"
        ) %>%
        pull(name)
    )
  )

table(short_term_subset$significant, short_term_subset$absent_in_long_term)

short_term_subset <- short_term_subset %>%
  mutate(
    state = case_when(
      significant == "Significant" & !absent_in_long_term ~ "Sig_in_long_term",
      significant == "Significant" & absent_in_long_term  ~ "Sig_not_in_long_term",
      significant == "Not significant" & !absent_in_long_term ~ "NS_in_long_term",
      TRUE ~ "NS_not_in_long_term"
    )
  )

genes_by_state <- short_term_subset %>%
  group_by(state) %>%
  summarise(
    genes = list(name),
    fc_values = list(FC_dep4),
    .groups = "drop"
  )

# Drug MoA in Resistant cells
(sig_long_term <- short_term_subset %>%
  filter(state == "Sig_in_long_term") %>%
  select(name, FC_dep4)%>% 
  arrange(desc(FC_dep4)))

# Drug MoA in Sensitive cells
(sig_not_long_term <- short_term_subset %>%
  filter(state == "Sig_not_in_long_term") %>%
  select(name, FC_dep4) %>% 
  arrange(desc(FC_dep4)))

# (ns_long_term <- short_term_subset %>%
#   filter(state == "NS_in_long_term") %>%
#   select(name, FC_dep4)%>% 
#   arrange(desc(FC_dep4)))
# 
# (ns_not_long_term <- short_term_subset %>%
#   filter(state == "NS_not_in_long_term") %>%
#   select(name, FC_dep4)%>% 
#   arrange(desc(FC_dep4)))
```

```{r}
library(dplyr)
library(ggplot2)
library(ggrepel)

df1 <- sig_long_term %>%
  arrange(FC_dep4) %>%
  mutate(
    gene = factor(name, levels = name),  # preserves ranking
    group = ifelse(FC_dep4 > 0.5, "> 0.5", "≤ 0.5")
  )

df2 <- sig_not_long_term %>%
  arrange(FC_dep4) %>%
  mutate(
    gene = factor(name, levels = name),  # preserves ranking
    group = ifelse(FC_dep4 > 0.5, "> 0.5", "≤ 0.5")
  )



r1_2 <- ggplot(df1, aes(x = gene, y = FC_dep4)) +
  
  # stems
  geom_segment(aes(xend = gene, y = 0, yend = FC_dep4, color = group),
               linewidth = 0.6) +
  
  # zero reference line
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  
  # bubbles
  geom_point(aes(color = group), size = 3) +
  
  scale_color_manual(values = c("> 0.5" = "#EA4335", "≤ 0.5" = "#4285F4")) +
  
  coord_flip() +   # 🔥 flip AFTER ordering
  
  theme_minimal(base_size = 12) +
  
  theme(
    axis.text.y = element_text(size = 10),  # gene labels now on y-axis
    axis.ticks = element_blank(),
    
    panel.grid = element_blank(),
    
    panel.border = element_rect(
      colour = "black",
      fill = NA,
      linewidth = 0.8
    ),
    
    legend.position = "bottom"
  ) +
  
  labs(
    x = NULL,
    y = NULL,
    color = "Long-term expression proteomics log2FC",
    title = "Long-term \nretained targets"
  )

r3_4 <- ggplot(df2, aes(x = gene, y = FC_dep4)) +
  
  # stems
  geom_segment(aes(xend = gene, y = 0, yend = FC_dep4, color = group),
               linewidth = 0.6) +
  
  # zero reference line
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  
  # bubbles
  geom_point(aes(color = group), size = 3) +
  
  scale_color_manual(values = c("> 0.5" = "#EA4335", "≤ 0.5" = "#4285F4")) +
  
  coord_flip() +   # 🔥 flip AFTER ordering
  
  theme_minimal(base_size = 12) +
  
  theme(
    axis.text.y = element_text(size = 10),  # gene labels now on y-axis
    axis.ticks = element_blank(),
    
    panel.grid = element_blank(),
    
    panel.border = element_rect(
      colour = "black",
      fill = NA,
      linewidth = 0.8
    ),
    
    legend.position = "bottom"
  ) +
  
  labs(
    x = NULL,
    y = NULL,
    color = "Long-term expression proteomics log2FC",
    title = "Long-term \nlost targets"
  )
```


```{r}
library(patchwork)

combined_plot <- (
  (r3_4  + r1_2) ) +
  plot_layout(guides = "collect") &
  theme(
    plot.title = element_text(face = "bold"),
    legend.position = "bottom",
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 9),
    legend.key.size = unit(0.6, "lines")
  )

combined_plot
```

