---
title: "VenAza_Feb27"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M')`"
output:
  pdf_document:
    latex_engine: lualatex
    keep_tex: true
    df_print: tibble
  html_document: default
---

## Load Library
```{r, eval=FALSE, echo =FALSE, message=FALSE, warning=FALSE}
library(rpx)
#library(msdata)
library(Spectra)
library(tidyverse)
#library(cleaver)
#library(MSnbase)
library(tidyverse)
library(mzR)
library(PSMatch)
library(pheatmap)
library(QFeatures)
#library(sageR)
library(SummarizedExperiment)
library(factoextra)
library(limma)
library(DEP)
library(ggrepel)
library(data.table)

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("rpx")
```

## 1. Read evidence.txt
```{r}
library(data.table)
## Read raw peptide TSV
evidence <- read_tsv("/Users/nvd448/Documents/VenAza Project/evidence.txt", show_col_types = FALSE)

setDT(evidence)
```

## 2. Identify TMT reporter intensity columns and Mapping metadata
```{r}
evidence <- evidence[(is.na(Reverse) | Reverse != "+") & (is.na(`Potential contaminant`) | `Potential contaminant` != "+")]
tmt_cols <- grep("^Reporter intensity corrected", colnames(evidence), value = TRUE)
evidence <- evidence[rowSums(as.matrix(evidence[, tmt_cols, with = FALSE])) > 0, ]

library (readxl)
map <- read_excel("~/Documents/VenAza Project/5LABELS_CoPISA_E4.xlsx", sheet = "Sheet1" )

```

## 3. Build assay matrices
```{r}

library(data.table)
library(dplyr)
library(stringr)
library(tibble)
library(QFeatures)
library(SummarizedExperiment)

# ---------------------
# Step 1: Filter evidence
# ---------------------
evidence <- evidence[(is.na(Reverse) | Reverse != "+") &
                     (is.na(`Potential contaminant`) | `Potential contaminant` != "+")]

tmt_cols <- grep("^Reporter intensity corrected", colnames(evidence), value = TRUE)

evidence <- evidence[
  rowSums(as.matrix(evidence[, tmt_cols, with = FALSE])) > 0
]

# ---------------------
# Step 2: Prepare sample annotation
# ---------------------
sample_annotation <- map %>%
  select(label, treatment, cell, `repeat`, sample, Experiment) %>%
  distinct() %>%
  filter(`Experiment` != 4) %>% 
  mutate(
    TMT_channel = str_replace(label, "TMTpro-", "TMT"),
    TMT_channel = gsub(" ", "", TMT_channel),
    Experiment = paste0("Kit", Experiment)
  )

# ---------------------
# Step 3: Define TMTpro 16plex channel order
# ---------------------
channel_map <- c(
  "TMT126","TMT127N","TMT127C","TMT128N",
  "TMT128C","TMT129N","TMT129C","TMT130N",
  "TMT130C","TMT131N","TMT131C","TMT132N",
  "TMT132C","TMT133N","TMT133C","TMT134N"
)

# ---------------------
# Step 4: Build QFeatures object
# ---------------------
qf <- QFeatures()

experiments <- unique(evidence$Experiment)

for (exp in experiments) {
  
  message("Processing: ", exp)
  
  exp_data <- evidence[Experiment == exp]
  
  tmt_cols_exp <- grep("^Reporter intensity corrected",
                       colnames(exp_data),
                       value = TRUE)
  
  # Safety check
  stopifnot(length(tmt_cols_exp) == 16)
  
  # ---- Match annotation to channels ----
  used_channels <- sample_annotation %>%
    filter(Experiment == exp) %>%
    pull(TMT_channel)
  
  keep_idx <- match(used_channels, channel_map)
  keep_idx <- keep_idx[!is.na(keep_idx)]
  
  # ---- Subset intensity columns ----
  exp_subset <- exp_data[, tmt_cols_exp[keep_idx], with = FALSE]
  
  mat <- as.matrix(exp_subset)
  mode(mat) <- "numeric"
  
  colnames(mat) <- paste0(exp, "_", used_channels)
  
  stopifnot(ncol(mat) == length(used_channels))
  
  # ---------------------
  # rowData
  # ---------------------
  row_data_exp <- DataFrame(
    Sequence        = exp_data$Sequence,
    ModifiedSeq     = exp_data$`Modified sequence`,
    Proteins        = exp_data$Proteins,
    LeadingProtein  = exp_data$`Leading proteins`,
    GeneNames       = exp_data$`Gene names`,
    Charge          = exp_data$Charge,
    PEP             = exp_data$PEP,
    Score           = exp_data$Score,
    RawFile         = exp_data$`Raw file`,
    Experiment      = exp_data$Experiment,
    Reverse         = exp_data$Reverse,
    Contaminant     = exp_data$`Potential contaminant`
  )
  
  rownames(row_data_exp) <- exp_data$id
  
  # ---------------------
  # colData
  # ---------------------
  col_data_exp <- sample_annotation %>%
    filter(Experiment == exp & TMT_channel %in% used_channels) %>%
    mutate(colname = paste0(Experiment, "_", TMT_channel)) %>%
    arrange(match(TMT_channel, used_channels)) %>%
    column_to_rownames("colname") %>%
    DataFrame()
  
  stopifnot(ncol(mat) == nrow(col_data_exp))
  
  # ---------------------
  # Create SummarizedExperiment
  # ---------------------
  se_exp <- SummarizedExperiment(
    assays  = list(intensity = mat),
    rowData = row_data_exp,
    colData = col_data_exp
  )
  
  qf[[exp]] <- se_exp
}

# ---------------------
# Inspect result
# ---------------------
qf
names(qf)
```

Inspect
```{r}
qf
summary(qf)
class(qf) 
names(qf)
colData(qf)
rowData(qf)
anyNA(assay(qf))
```

## 4. Normalize psm intensities

```{r}
library(matrixStats)

psm_sets <- names(qf)

for (psm_name in psm_sets) {
  
  message("Processing normalization: ", psm_name)
  
  # -------------------------
  # 1) Extract raw assay
  # -------------------------
  raw_assay <- assay(qf[[psm_name]])
  raw_assay <- as.matrix(raw_assay)
  
  # -------------------------
  # 2) Fraction of column total
  # -------------------------
  frac_assay <- sweep(raw_assay, 2,
                      colSums(raw_assay, na.rm = TRUE),
                      FUN = "/")
  
  scaled_assay <- frac_assay * 1e6
  
  frac_name <- paste0(psm_name, "_frac")
  
  qf[[frac_name]] <- SummarizedExperiment(
    assays  = list(intensity = scaled_assay),
    rowData = rowData(qf[[psm_name]]),
    colData = colData(qf[[psm_name]])
  )
  
  qf <- addAssayLinkOneToOne(qf,
                             from = psm_name,
                             to   = frac_name)
  
  # -------------------------
  # 3) Add pseudocount
  # -------------------------
  frac_pc <- scaled_assay + 1e-6
  
  frac_pc_name <- paste0(psm_name, "_frac_pc")
  
  qf[[frac_pc_name]] <- SummarizedExperiment(
    assays  = list(intensity = frac_pc),
    rowData = rowData(qf[[frac_name]]),
    colData = colData(qf[[frac_name]])
  )
  
  qf <- addAssayLinkOneToOne(qf,
                             from = frac_name,
                             to   = frac_pc_name)
  
  # -------------------------
  # 4) log2 transform
  # -------------------------
  log_name <- paste0(psm_name, "_log")
  
  qf <- logTransform(
    qf,
    i    = frac_pc_name,
    name = log_name
  )
  
  # -------------------------
  # 5) Quantile normalize
  # -------------------------
  norm_name <- paste0(psm_name, "_norm")
  
  qf <- normalize(
    qf,
    i      = log_name,
    name   = norm_name,
    method = "quantiles"
  )
}

```

## 5. Aggregation
```{r}
norm_sets <- grep("_norm$", names(qf), value = TRUE)

for (norm_name in norm_sets) {
  
  message("Aggregating: ", norm_name)
  
  # -------------------------
  # PSM → Peptide
  # -------------------------
  pep_name <- paste0(norm_name, "_peptide")
  
  qf <- aggregateFeatures(
    object = qf,
    i      = norm_name,
    fcol   = "Sequence",
    name   = pep_name,
    fun    = function(x)
      matrixStats::colMedians(x, na.rm = TRUE)
  )
  
  # -------------------------
  # Peptide → Protein
  # -------------------------
  prot_name <- paste0(norm_name, "_protein")
  
  qf <- aggregateFeatures(
    object = qf,
    i      = pep_name,
    fcol   = "Proteins",
    name   = prot_name,
    fun    = function(x)
      matrixStats::colMedians(x, na.rm = TRUE)
  )
}
```

Inspect
```{r}
qf
summary(qf)
class(qf)
names(qf)
colData(qf)
rowData(qf)
anyNA(assay(qf))
library(dplyr)
library(purrr)

qfInfo <- tibble(
  assay  = names(qf),
  n_rows = map_int(qf, nrow),
  n_cols = map_int(qf, ncol)
) %>%
  arrange(assay)

qfInfo

plot(qf)
```
```{r}
library(QFeatures)
library(dplyr)
library(purrr)
library(ggplot2)

# Select assays
protein_assays <- grep("^Kit4_.*_norm_protein$", names(qf), value = TRUE)
peptide_assays <- grep("^Kit4_.*_norm_peptide$", names(qf), value = TRUE)

# Friendly labels
condition_labels <- c(
  "Kit4_2" = "Short-term living cell CoPISA",
  "Kit4_3" = "Long-term living cell CoPISA",
  "Kit4_4" = "Short-term Cell lysate CoPISA",
  "Kit4_5" = "Long-term expression proteomics"
)

# Count identified features in each assay
count_identified <- function(assays, type) {
  tibble(
    assay = assays,
    identified = map_int(assays, ~{
      mat <- assay(qf, .x)
      sum(rowSums(!is.na(mat)) > 0)
    }),
    Type = type
  ) %>%
    mutate(
      Condition = condition_labels[sub("_norm_(protein|peptide)$", "", assay)]
    )
}

protein_counts <- count_identified(protein_assays, "Protein")
peptide_counts <- count_identified(peptide_assays, "Peptide")

# Total unique proteins
total_proteins <- length(unique(unlist(
  lapply(protein_assays, function(a) {
    mat <- assay(qf, a)
    rownames(mat)[rowSums(!is.na(mat)) > 0]
  })
)))

# Total unique peptides
total_peptides <- length(unique(unlist(
  lapply(peptide_assays, function(a) {
    mat <- assay(qf, a)
    rownames(mat)[rowSums(!is.na(mat)) > 0]
  })
)))

# Add totals
total_df <- tibble(
  Condition = "Total unique",
  identified = c(total_proteins, total_peptides),
  Type = c("Protein", "Peptide")
)

plot_df <- bind_rows(
  protein_counts %>% select(Condition, identified, Type),
  peptide_counts %>% select(Condition, identified, Type),
  total_df
)
plot_df$Condition <- factor(
  plot_df$Condition,
  levels = c(
    "Short-term living cell CoPISA",
    "Long-term living cell CoPISA",
    "Short-term Cell lysate CoPISA",
    "Long-term expression proteomics",
    "Total unique"
  )
)

plot_df <- plot_df %>%
  arrange(Condition, Type) %>%
  mutate(
    group = as.numeric(Condition),
    x = ifelse(
      Type == "Protein",
      group - 0.15,
      group + 0.15
    )
  )
# Preserve order
plot_df$Condition <- factor(
  plot_df$Condition,
  levels = c(
    "Short-term living cell CoPISA",
    "Long-term living cell CoPISA",
    "Short-term Cell lysate CoPISA",
    "Long-term expression proteomics",
    "Total unique"
  )
)

# Plot
library(ggplot2)

ggplot(plot_df, aes(x = x, y = identified, colour = Type)) +

  geom_segment(aes(xend = x,
                   y = 0,
                   yend = identified),
               linewidth = 1) +

  geom_point(size = 4) +

  geom_text(aes(label = identified),
            vjust = -0.7,
            size = 3.8,
            show.legend = FALSE) +

  scale_x_continuous(
    breaks = 1:5,
    labels = c(
      "Short-term living cell CoPISA",
      "Long-term living cell CoPISA",
      "Short-term Cell lysate CoPISA",
      "Long-term expression proteomics",
      "Total unique"
    )
  ) +

  scale_colour_manual(values = c(
    Protein = "#0072B2",
    Peptide = "#D55E00"
  )) +

  expand_limits(y = max(plot_df$identified) * 1.08) +

  labs(
    x = NULL,
    y = "Number of identified features",
    colour = NULL
  ) +

  theme_classic(base_size = 14) +
  theme(
    legend.position = "right",
    axis.text.x = element_text(size = 11)
  )+ coord_flip()
```

## 6. Differential analysis

```{r, eval=FALSE, echo =FALSE, message=FALSE, warning=FALSE}
library(DEP)
library(dplyr)
library(tidyr)
library(S4Vectors)

protein_sets <- grep("_norm_protein$", names(qf), value = TRUE)

dep_results <- list()

for (prot_name in protein_sets) {
  
  message("Running DEP for: ", prot_name)
  
  protein_se <- qf[[prot_name]]
  
  # -------------------------
  # Build colData
  # -------------------------
  
  col_df <- as.data.frame(colData(protein_se))
  col_df$label     <- rownames(col_df)
  col_df$condition <- col_df$treatment
  col_df$replicate <- col_df$`repeat`
  
  # -------------------------
  # Build rowData (KEEP GeneNames)
  # -------------------------
  
  row_df <- as.data.frame(rowData(protein_se))
  
  # Make sure GeneNames column exists
  if (!"GeneNames" %in% colnames(row_df)) {
    stop("GeneNames column not found in rowData(protein_se)")
  }
  
  row_df$name  <- row_df$Proteins
  row_df$ID    <- paste0(row_df$Proteins, "_", seq_len(nrow(row_df)))
  row_df$gene  <- row_df$GeneNames   # 👈 IMPORTANT ADDITION
  
  assay_mat <- assay(protein_se)
  row_df <- cbind(row_df, assay_mat)
  
  columns <- (ncol(row_df) - ncol(assay_mat) + 1):ncol(row_df)
  
  # -------------------------
  # Make DEP object
  # -------------------------
  
  dep <- make_se(row_df, columns, col_df)
  
  # Impute
  set.seed(123456)
  dep <- DEP::impute(dep, fun = "MinProb", q = 0.01)
  
  # Remove Pool
  dep <- dep[, dep$condition != "Pool"]
  
  # Control
  dep <- test_diff(
    dep,
    type    = "control",
    control = "C"
  )
  
  dep <- add_rejections(
    dep,
    alpha = 0.05,
    lfc   = 0
  )
  
  # Store results (NOW includes gene column)
  res <- get_results(dep)

  
# -------------------------
# Normalize Kit4.3 vs Kit4.5 (ALL contrasts)
# -------------------------

if (prot_name == "Kit4_3_norm_protein") {

  kit45 <- dep_results[["Kit4_5_norm_protein"]]

  if (!is.null(kit45)) {

    kit45_fc <- kit45 %>%
      dplyr::select(
        name,
        A_vs_C_ratio_45  = A_vs_C_ratio,
        V_vs_C_ratio_45  = V_vs_C_ratio,
        VA_vs_C_ratio_45 = VA_vs_C_ratio
      )

    res <- res %>%
      dplyr::left_join(kit45_fc, by = "name") %>%
      dplyr::mutate(
        A_vs_C_ratio  = A_vs_C_ratio  - A_vs_C_ratio_45,
        V_vs_C_ratio  = V_vs_C_ratio  - V_vs_C_ratio_45,
        VA_vs_C_ratio = VA_vs_C_ratio - VA_vs_C_ratio_45
      ) %>%
      dplyr::select(
        -A_vs_C_ratio_45,
        -V_vs_C_ratio_45,
        -VA_vs_C_ratio_45
      )
  }
}
# Extract gene names from DEP object
gene_df <- as.data.frame(rowData(dep))[, c("name", "gene")]

# Merge gene names back into results
res <- left_join(res, gene_df, by = "name")

dep_results[[prot_name]] <- res
}
```

```{r}
## Checking different thresholds
alpha_vals <- c(0.01, 0.05, 0.1)
lfc_vals   <- c(0, 0.5, 1)

sig_summary <- list()

for (prot_name in protein_sets) {

  message("Running DEP for: ", prot_name)
  
  protein_se <- qf[[prot_name]]
  
  col_df <- as.data.frame(colData(protein_se))
  col_df$label     <- rownames(col_df)
  col_df$condition <- col_df$treatment
  col_df$replicate <- col_df$`repeat`
  
  row_df <- as.data.frame(rowData(protein_se))
  row_df$name  <- row_df$Proteins
  row_df$ID    <- paste0(row_df$Proteins, "_", seq_len(nrow(row_df)))
  row_df$gene  <- row_df$GeneNames
  
  assay_mat <- assay(protein_se)
  row_df <- cbind(row_df, assay_mat)
  columns <- (ncol(row_df) - ncol(assay_mat) + 1):ncol(row_df)
  
  dep <- make_se(row_df, columns, col_df)
  dep <- impute(dep, fun ="MinProb", q=0.01)
  dep <- dep[, dep$condition != "Pool"]
  
  dep <- test_diff(dep, type="control", control="C")
  
  # Test multiple alpha / lfc combinations
  for(a in alpha_vals){
    for(l in lfc_vals){
      
      dep_tmp <- add_rejections(dep, alpha=a, lfc=l)
      
      res <- get_results(dep_tmp)
      
      n_sig <- sum(res$significant, na.rm=TRUE)
      
      sig_summary[[length(sig_summary)+1]] <- data.frame(
        dataset = prot_name,
        alpha   = a,
        lfc     = l,
        significant_proteins = n_sig
      )
    }
  }
}

sig_summary <- bind_rows(sig_summary)
sig_summary
ggplot(sig_summary, aes(x=lfc, y=significant_proteins, color=factor(alpha))) +
  geom_line() +
  geom_point(size=3) +
  facet_wrap(~dataset) +
  labs(
    color="alpha",
    x="log2FC threshold",
    y="Number of significant proteins"
  ) +
  theme_bw()
```
### Permutation test

```{r}
run_hybrid_permutation <- function(dep,
                                   control = "C",
                                   alpha = 0.05,
                                   lfc = 0,
                                   max_exact = 50000,
                                   n_mc = 10,
                                   seed = 1) {

  set.seed(seed)

  # -----------------------
  # ORIGINAL LABELS
  # -----------------------
  original_labels <- dep$condition
  n <- length(original_labels)

  # -----------------------
  # GENERATE UNIQUE PERMUTATIONS (EXACT MODE)
  # -----------------------
  perm_list <- gtools::permutations(
    n = n,
    r = n,
    v = original_labels
  )

  perm_list <- unique(perm_list)

  n_perm_exact <- nrow(perm_list)

  use_exact <- n_perm_exact <= max_exact

  # -----------------------
  # OBSERVED RESULT
  # -----------------------
  dep_obs <- test_diff(dep, type = "control", control = control)
  dep_obs <- add_rejections(dep_obs, alpha = alpha, lfc = lfc)

  obs_sig <- sum(get_results(dep_obs)$significant, na.rm = TRUE)

  # -----------------------
  # STORAGE
  # -----------------------
  perm_sig <- NULL

  # =========================================================
  # MODE 1: EXACT PERMUTATION
  # =========================================================
  if (use_exact) {

    message("Using EXACT permutation: ", n_perm_exact, " permutations")

    perm_sig <- numeric(n_perm_exact)

    for (i in seq_len(n_perm_exact)) {

      dep_perm <- dep
      dep_perm$condition <- perm_list[i, ]

      dep_perm <- test_diff(dep_perm, type = "control", control = control)
      dep_perm <- add_rejections(dep_perm, alpha = alpha, lfc = lfc)

      perm_sig[i] <- sum(get_results(dep_perm)$significant, na.rm = TRUE)
    }

    n_perm_used <- n_perm_exact
  }

  # =========================================================
  # MODE 2: MONTE CARLO PERMUTATION
  # =========================================================
  else {

    message("Using MONTE CARLO permutation: ", n_mc, " iterations")

    perm_sig <- numeric(n_mc)

    for (i in seq_len(n_mc)) {

      dep_perm <- dep

      # shuffle labels (correct null model)
      dep_perm$condition <- sample(original_labels)

      dep_perm <- test_diff(dep_perm, type = "control", control = control)
      dep_perm <- add_rejections(dep_perm, alpha = alpha, lfc = lfc)

      perm_sig[i] <- sum(get_results(dep_perm)$significant, na.rm = TRUE)
    }

    n_perm_used <- n_mc
  }

  # -----------------------
  # P-VALUE
  # -----------------------
  p_global <- (sum(perm_sig >= obs_sig) + 1) / (n_perm_used + 1)

  # -----------------------
  # OUTPUT
  # -----------------------
  list(
    observed_significant = obs_sig,
    permutation_distribution = perm_sig,
    global_p_value = p_global,
    n_permutations_used = n_perm_used,
    mode = ifelse(use_exact, "exact", "monte_carlo")
  )
}
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
perm_results <- list()

set.seed(123)

for (prot_name in protein_sets) {

  message("\nRunning permutation DEP for: ", prot_name)

  protein_se <- qf[[prot_name]]

  # Ask user for parameters before each run
  alpha <- as.numeric(readline(prompt = "Enter alpha (e.g. 0.05, 0.001): "))
  lfc   <- as.numeric(readline(prompt = "Enter lfc (e.g. 0, 1): "))

  perm_results[[prot_name]] <- run_dep_permutation(
    protein_se = protein_se,
    n_perm = 10000,
    control = "C",
    alpha = alpha,
    lfc = lfc,
    seed = 123
  )
}

summary_table <- do.call(rbind, lapply(names(perm_results), function(nm) {

  res <- perm_results[[nm]]

  data.frame(
    dataset = nm,
    observed_sig = res$observed_significant,
    global_p = res$global_p_value
  )
}))

```

## 7. Upset plot 1


```{r}
library(dplyr)
library(UpSetR)
library(viridisLite)

vcols <- viridis(8)

# Google colors for sets
treatment_colors <- c("A"=vcols[2],"V"=vcols[4],"VA"=vcols[6])

make_upset <- function(res, prefix) {

  upset_df <- res %>%
    dplyr::select(name, A_vs_C_significant, V_vs_C_significant, VA_vs_C_significant) %>%
    dplyr::mutate(
      A = as.integer(A_vs_C_significant),
      V = as.integer(V_vs_C_significant),
      VA = as.integer(VA_vs_C_significant)
    ) %>%
    dplyr::select(name, A, V, VA)

  upset_df <- upset_df[rowSums(upset_df[, -1]) > 0, ]

  if(nrow(upset_df) == 0) {
    message("No significant proteins to plot for ", prefix)
    return(NULL)
  }

  set_order <- c("V","A","VA")

  UpSetR::upset(
    upset_df,
     # Set structure
    sets = set_order,
    nsets = 3,
    nintersects = NA,
    order.by = "freq",
    decreasing = TRUE,
    main.bar.color = vcols[5],
    sets.bar.color = treatment_colors[set_order],  # match order
    # Matrix appearance
    matrix.color = vcols[3],
    point.size = 4,
    line.size = 1.2,
    mb.ratio = c(0.65, 0.35),
    # Text scaling
    text.scale = c(1.4, 1.6, 1, 1.2, 1.6, 1.4),
    # Intersection numbers
    show.numbers = "yes",
    number.angles = 30,
    # Background shading
    shade.color = vcols[1],
    shade.alpha = 0.2,
    mainbar.y.label = "Number of proteins",
    sets.x.label = "Proteins per treatment",
    keep.order = TRUE
  )
}
    
# kit_titles <- c(
#   "Kit4_2_norm_protein" = "Short-term living cell CoPISA",
#   "Kit4_3_norm_protein" = "Long-term living cell CoPISA",
#   "Kit4_4_norm_protein" = "Short-term Cell lysate CoPISA",
#   "Kit4_5_norm_protein" = "Long-term expression proteomics"
# )


up1 <- make_upset(dep_results[[1]], "Kit4_2_norm_protein")
up2 <- make_upset(dep_results[[2]], "Kit4_3_norm_protein")
up3 <- make_upset(dep_results[[3]], "Kit4_4_norm_protein")
up4 <- make_upset(dep_results[[4]], "Kit4_5_norm_protein")
up2
up3
up1
up4


# Save up1 as PNG
png("up1.png", width = 10, height = 10, units = "cm", res = 300)
print(up1)  # must use print() for ComplexUpset
dev.off()

# Repeat for other plots
png("up2.png", width = 10, height = 10, units = "cm", res = 300)
print(up2)
dev.off()

png("up3.png", width = 10, height = 10, units = "cm", res = 300)
print(up3)
dev.off()

png("up4.png", width = 10, height = 10, units = "cm", res = 300)
print(up4)
dev.off()
```

### Hypergeometric overlap + VA-specific fraction

```{r}
library(dplyr)

hypergeom_overlap <- function(set1, set2, universe) {

  # observed overlap
  k <- length(intersect(set1, set2))

  # size of sets
  K <- length(set1)
  n <- length(set2)

  # universe size
  N <- length(universe)

  # hypergeometric test
  pval <- phyper(
    k - 1,
    K,
    N - K,
    n,
    lower.tail = FALSE
  )

  expected <- (K * n) / N

  data.frame(
    set1_size = K,
    set2_size = n,
    overlap = k,
    expected_overlap = round(expected, 2),
    enrichment = round(k / expected, 2),
    p_value = pval
  )
}


analyze_VA_specificity <- function(res, dataset_name) {


  # All quantified proteins = universe
  universe <- unique(res$name)


  # Significant protein sets
  A <- res %>%
    filter(A_vs_C_significant) %>%
    pull(name)

  V <- res %>%
    filter(V_vs_C_significant) %>%
    pull(name)

  VA <- res %>%
    filter(VA_vs_C_significant) %>%
    pull(name)



  # ----------------------------
  # Hypergeometric tests
  # ----------------------------

  VA_vs_A <- hypergeom_overlap(
    set1 = VA,
    set2 = A,
    universe = universe
  ) %>%
    mutate(
      comparison = "VA vs A",
      dataset = dataset_name
    )


  VA_vs_V <- hypergeom_overlap(
    set1 = VA,
    set2 = V,
    universe = universe
  ) %>%
    mutate(
      comparison = "VA vs V",
      dataset = dataset_name
    )


  overlap_stats <- bind_rows(
    VA_vs_A,
    VA_vs_V
  )



  # ----------------------------
  # VA-specific proteins
  # ----------------------------

  VA_unique <- setdiff(
    VA,
    union(A,V)
  )


  VA_fraction <- data.frame(
    dataset = dataset_name,
    VA_total = length(VA),
    VA_unique = length(VA_unique),
    VA_specific_fraction =
      length(VA_unique) / length(VA)
  )


  list(
    overlap = overlap_stats,
    VA_specific = VA_fraction,
    VA_unique_proteins = VA_unique
  )

}
```

```{r}
results_statistics <- lapply(
  names(dep_results),
  function(x) {

    analyze_VA_specificity(
      dep_results[[x]],
      x
    )

  }
)

names(results_statistics) <- names(dep_results)

hypergeom_table <- bind_rows(
  lapply(results_statistics, function(x)
    x$overlap)
)

hypergeom_table

VA_specific_table <- bind_rows(
  lapply(results_statistics, function(x)
    x$VA_specific)
)

VA_specific_table

hypergeom_table <- hypergeom_table %>%
  mutate(
    FDR = p.adjust(
      p_value,
      method = "BH"
    )
  )

hypergeom_table
```



## 8. Upset plot 2

```{r}
library(dplyr)
library(tidyr)
library(UpSetR)
library(viridis)
library(purrr)  # for reduce()

# ---- 1. Prepare combined VA significant dataframe ----
upset_combined <- lapply(names(dep_results), function(nm) {
  res <- dep_results[[nm]]
  
  res %>%
    select(name = gene, VA_vs_C_significant) %>%
    mutate(!!nm := as.integer(VA_vs_C_significant)) %>%
    select(-VA_vs_C_significant)
}) %>%
  purrr::reduce(full_join, by = "name")  # combine all 4 datasets by protein name

# Replace NAs with 0 (not significant)
upset_combined[is.na(upset_combined)] <- 0

# ---- 2. Map kit names to descriptive titles ----
kit_titles <- c(
  "Kit4_2_norm_protein" = "Short-term living cell CoPISA",
  "Kit4_3_norm_protein" = "Long-term living cell CoPISA",
  "Kit4_4_norm_protein" = "Short-term Cell lysate CoPISA",
  "Kit4_5_norm_protein" = "Long-term expression proteomics"
)

# Replace column names with descriptive names
colnames(upset_combined) <- c("name", kit_titles[colnames(upset_combined)[-1]])

# ---- 3. Define set order and colors ----
set_order <- kit_titles  # use descriptive names

pcols <- plasma(6)
treatment_colors <- c(
  "Short-term living cell CoPISA"    = pcols[2],
  "Long-term living cell CoPISA"     = pcols[3],
  "Short-term Cell lysate CoPISA"    = pcols[4],
  "Long-term expression proteomics"  = pcols[5]
)

upset_combined_VA <- upset_combined

# ---- 4. Create UpSet plot ----

upset_VA <- UpSetR::upset(
  upset_combined,
  
  sets = set_order,
  nsets = length(set_order),
  nintersects = NA,
  order.by = "freq",
  decreasing = TRUE,
  
  main.bar.color = vcols[6],                     # intersection bars
  sets.bar.color = treatment_colors[set_order],  # treatment bars
  matrix.color = vcols[3],                       # intersection points
  point.size = 4,
  line.size = 1.2,
  mb.ratio = c(0.65, 0.35),
  
  text.scale = c(1.4, 1.6, 1, 1.2, 1.6, 1.4),
  
  show.numbers = "yes",
  number.angles = 30,
  
  shade.color = vcols[1],
  shade.alpha = 0.2,
  
  mainbar.y.label = "Number of proteins",
  sets.x.label = "Proteins per VA \ntreatment in all four samples",
  keep.order = TRUE
)

```

```{r}
# ---- 1. Prepare combined V significant dataframe ----
upset_combined <- lapply(names(dep_results), function(nm) {
  res <- dep_results[[nm]]
  
  res %>%
    select(name = gene, V_vs_C_significant) %>%
    mutate(!!nm := as.integer(V_vs_C_significant)) %>%
    select(-V_vs_C_significant)
}) %>%
  purrr::reduce(full_join, by = "name")  # combine all 4 datasets by protein name

# Replace NAs with 0 (not significant)
upset_combined[is.na(upset_combined)] <- 0

# ---- 2. Map kit names to descriptive titles ----
kit_titles <- c(
  "Kit4_2_norm_protein" = "Short-term living cell CoPISA",
  "Kit4_3_norm_protein" = "Long-term living cell CoPISA",
  "Kit4_4_norm_protein" = "Short-term Cell lysate CoPISA",
  "Kit4_5_norm_protein" = "Long-term expression proteomics"
)

# Replace column names with descriptive names
colnames(upset_combined) <- c("name", kit_titles[colnames(upset_combined)[-1]])

# ---- 3. Define set order and colors ----
set_order <- kit_titles  # use descriptive names

pcols <- plasma(6)
treatment_colors <- c(
  "Short-term living cell CoPISA"    = pcols[2],
  "Long-term living cell CoPISA"     = pcols[3],
  "Short-term Cell lysate CoPISA"    = pcols[4],
  "Long-term expression proteomics"  = pcols[5]
)

upset_combined_V <- upset_combined

# ---- 4. Create UpSet plot ----

upset_V <- UpSetR::upset(
  upset_combined,
  
  sets = set_order,
  nsets = length(set_order),
  nintersects = NA,
  order.by = "freq",
  decreasing = TRUE,
  
  main.bar.color = vcols[4],                     # intersection bars
  sets.bar.color = treatment_colors[set_order],  # treatment bars
  matrix.color = vcols[3],                       # intersection points
  point.size = 4,
  line.size = 1.2,
  mb.ratio = c(0.65, 0.35),
  
  text.scale = c(1.4, 1.6, 1, 1.2, 1.6, 1.4),
  
  show.numbers = "yes",
  number.angles = 30,
  
  shade.color = vcols[1],
  shade.alpha = 0.2,
  
  mainbar.y.label = "Number of proteins",
  sets.x.label = "Proteins per V \ntreatment in all four samples",
  keep.order = TRUE
)

```

```{r}
# ---- 1. Prepare combined A significant dataframe ----
upset_combined <- lapply(names(dep_results), function(nm) {
  res <- dep_results[[nm]]
  
  res %>%
    select(name = gene, A_vs_C_significant) %>%
    mutate(!!nm := as.integer(A_vs_C_significant)) %>%
    select(-A_vs_C_significant)
}) %>%
  purrr::reduce(full_join, by = "name")  # combine all 4 datasets by protein name

# Replace NAs with 0 (not significant)
upset_combined[is.na(upset_combined)] <- 0

# ---- 2. Map kit names to descriptive titles ----
kit_titles <- c(
  "Kit4_2_norm_protein" = "Short-term living cell CoPISA",
  "Kit4_3_norm_protein" = "Long-term living cell CoPISA",
  "Kit4_4_norm_protein" = "Short-term Cell lysate CoPISA",
  "Kit4_5_norm_protein" = "Long-term expression proteomics"
)

# Replace column names with descriptive names
colnames(upset_combined) <- c("name", kit_titles[colnames(upset_combined)[-1]])

# ---- 3. Define set order and colors ----
set_order <- kit_titles  # use descriptive names

pcols <- plasma(6)
treatment_colors <- c(
  "Short-term living cell CoPISA"    = pcols[2],
  "Long-term living cell CoPISA"     = pcols[3],
  "Short-term Cell lysate CoPISA"    = pcols[4],
  "Long-term expression proteomics"  = pcols[5]
)

upset_combined_A <- upset_combined

# ---- 4. Create UpSet plot ----

upset_A <- UpSetR::upset(
  upset_combined,
  
  sets = set_order,
  nsets = length(set_order),
  nintersects = NA,
  order.by = "freq",
  decreasing = TRUE,
  
  main.bar.color = vcols[2],                     # intersection bars
  sets.bar.color = treatment_colors[set_order],  # treatment bars
  matrix.color = vcols[3],                       # intersection points
  point.size = 4,
  line.size = 1.2,
  mb.ratio = c(0.65, 0.35),
  
  text.scale = c(1.4, 1.6, 1, 1.2, 1.6, 1.4),
  
  show.numbers = "yes",
  number.angles = 30,
  
  shade.color = vcols[1],
  shade.alpha = 0.2,
  
  mainbar.y.label = "Number of proteins",
  sets.x.label = "Proteins per A \ntreatment in all four samples",
  keep.order = TRUE
)

```


```{r}
png("upset_VA.png", width = 15, height = 10, units = "cm", res = 300)
upset_VA
dev.off()

png("upset_V.png", width = 15, height = 10, units = "cm", res = 300)
upset_V
dev.off()

png("upset_A.png", width = 15, height = 10, units = "cm", res = 300)
upset_A
dev.off()

upset_VA
upset_V
upset_A
```

### Hypergeometric overlap test

```{r}
pairwise_hyper <- function(upset_df){

    kits <- names(upset_df)[-1]

    universe <- unique(upset_df$name)

    res <- list()

    counter <- 1

    for(i in 1:(length(kits)-1)){

        for(j in (i+1):length(kits)){

            set1 <- upset_df$name[upset_df[[kits[i]]]==1]
            set2 <- upset_df$name[upset_df[[kits[j]]]==1]

            N <- length(universe)
            K <- length(set1)
            n <- length(set2)

            k <- length(intersect(set1,set2))

            expected <- K*n/N

            p <- phyper(
                k-1,
                K,
                N-K,
                n,
                lower.tail=FALSE
            )

            res[[counter]] <- data.frame(
                comparison=paste(kits[i],"vs",kits[j]),
                set1=K,
                set2=n,
                overlap=k,
                expected=expected,
                enrichment=k/expected,
                p_value=p
            )

            counter <- counter+1

        }

    }

    bind_rows(res) |>
        mutate(
            FDR=p.adjust(p_value,"BH")
        )

}

VA_overlap_stats <- pairwise_hyper(upset_combined_VA)

V_overlap_stats <- pairwise_hyper(upset_combined_V)

A_overlap_stats <- pairwise_hyper(upset_combined_A)
```


### Extract gene name of common sets
```{r}
# library(dplyr)
# 
# set_cols <- colnames(upset_combined)[-1]
# n_sets <- length(set_cols)

# Create combinations
comb_df <- upset_combined %>%
  mutate(
    combination = apply(select(., all_of(set_cols)), 1, function(x) {
      paste(set_cols[x == 1], collapse = " & ")
    }),
    n_members = rowSums(select(., all_of(set_cols)))
  ) %>%
  filter(combination != "")


#---- OPTION 2 (alternative): only shared sets (>=2) ----
(intersection_labels <- comb_df %>%
  filter(n_members >= 2) %>%
  group_by(combination) %>%
  summarise(genes = paste(name, collapse = ", "), .groups = "drop")
)

```



## 9. Volcano plot
```{r}
# library(dplyr)
# library(tidyr)
# library(ggplot2)
# library(viridis)
# library(ggrepel)

make_volcano <- function(res, prefix) {
  
  pval_thresh <- 0.05
  lfc_thresh  <- 1
  
  # Kit name mapping
  kit_labels <- c(
    "Kit4_2_norm_protein" = "Intact_short_term",
    "Kit4_3_norm_protein" = "Intact_long_term",
    "Kit4_4_norm_protein" = "Extract_short_term",
    "Kit4_5_norm_protein" = "Expression_long_term"
  )
  
  plot_name <- kit_labels[prefix]
  if (is.na(plot_name)) plot_name <- prefix
  
  # Prepare long format
  long_df <<- res %>%
  select(name, gene,
         A_vs_C_ratio, A_vs_C_p.val, A_vs_C_p.adj,
         V_vs_C_ratio, V_vs_C_p.val, V_vs_C_p.adj,
         VA_vs_C_ratio, VA_vs_C_p.val, VA_vs_C_p.adj) %>%

  pivot_longer(
    cols = -c(name, gene),
    names_to = c("Treatment", ".value"),
    names_pattern = "(.*)_vs_C_(ratio|p\\.val|p\\.adj)"
  ) %>%

  mutate(
    log2FC = ratio,
    pval   = p.val,
    padj   = p.adj   # 🔥 now this exists
  ) %>%

  mutate(
    Treatment = recode(Treatment,
                       "A"  = "Azacytidine",
                       "VA" = "Azacytidine + Venetoclax",
                       "V"  = "Venetoclax"),
    negLogP = -log10(pval),
    Significant = (pval < pval_thresh & abs(log2FC) > lfc_thresh),
    gene = ifelse(is.na(gene) | gene == "", name, gene),
    gene = sub(";.*", "", gene)
  )
  
  # Determine shared vs unique (ONLY among significant proteins)
  shared_info <- long_df %>%
    filter(Significant) %>%
    group_by(name) %>%
    summarise(n_sig = n(), .groups = "drop")
  
  long_df <- long_df %>%
    left_join(shared_info, by = "name") %>%
    mutate(
      Category = case_when(
        Significant & n_sig > 1  ~ "Shared significant protein",
        Significant & n_sig == 1 ~ "Unique significant protein",
        TRUE ~ "Not significant"
      )
    )
  
  # 🔥 Select top 5 significant proteins per treatment (smallest p-value)
  label_df <- long_df %>%
    filter(Significant) %>%
    group_by(Treatment) %>%
    arrange(pval, .by_group = TRUE) %>%
    slice_head(n = 5) %>%
    ungroup()
  
  # 🔥 MODIFIED: define custom viridis colors
  library(viridis)
  vcols <- viridis(8)
  treatment_colors <- c(
    "Azacytidine"             = vcols[2],
    "Venetoclax"              = vcols[4],
    "Azacytidine + Venetoclax"= vcols[6]
  )
  
  p <- ggplot(long_df, aes(x = log2FC, y = negLogP)) +
    geom_point(aes(color = Treatment,
                   fill  = Category,
                   alpha = Significant),
               shape = 21,
               size = 1.2,
               stroke = 1
               ) +
  
    scale_alpha_manual(values = c(
      "TRUE"  = 1,
      "FALSE" = 0.15
    ), guide = "none") +
    
    geom_text_repel(
                data = label_df,
                aes(label = gene, color = Treatment),
                size = 3,
                box.padding = 0.3,
                max.overlaps = 20,
                force = 1,
                segment.size = 0.3,
                segment.alpha = 0.5
    ) +
    
    geom_vline(xintercept = c(-lfc_thresh, lfc_thresh),
               linetype = "dashed") +
    geom_hline(yintercept = -log10(pval_thresh),
               linetype = "dashed") +
    
    # 🔥 MODIFIED: use manual color assignment instead of viridis_d()
    scale_color_manual(values = treatment_colors) +
    
    scale_fill_manual(values = c(
      "Shared significant protein" = "black",
      "Unique significant protein" = "gray60",
      "Not significant" = "white"
    )) +
    
    theme_minimal(base_size = 13) +
    labs(
      title = paste("Volcano Plot -", plot_name),
      x = "log2 Fold Change",
      y = "-log10(p-value)"
    ) +
    theme(
      plot.title = element_text(face = "bold"),
      legend.position = "bottom",
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    )
  
  return(p)
}


```



```{r}
#library(patchwork)

# ---- 1. Define consistent axis limits ----
all_long <- lapply(names(dep_results), function(nm) {
  res <- dep_results[[nm]]
  
  res %>%
    select(A_vs_C_ratio, A_vs_C_p.val,
           V_vs_C_ratio, V_vs_C_p.val,
           VA_vs_C_ratio, VA_vs_C_p.val) %>%
    pivot_longer(everything(),
                 names_to = c("Treatment", ".value"),
                 names_pattern = "(.*)_vs_C_(.*)") %>%
    dplyr::rename(log2FC = ratio,
           pval   = p.val) %>%
    mutate(negLogP = -log10(pval))
})

all_long <- bind_rows(all_long)

x_limits <- range(all_long$log2FC, na.rm = TRUE)
y_limits <- range(all_long$negLogP, na.rm = TRUE)

# ---- 2. Custom panel titles ----
custom_titles <- c(
  "Kit4_2_norm_protein" = "Short-term living cell CoPISA",
  "Kit4_3_norm_protein" = "Long-term living cell CoPISA",
  "Kit4_4_norm_protein" = "Short-term Cell lysate CoPISA",
  "Kit4_5_norm_protein" = "Long-term expression protoemics"
)

# ---- 3. Create plots with modifications ----
plots <- lapply(names(dep_results), function(nm) {
  
  p <- make_volcano(dep_results[[nm]], nm) +
    coord_cartesian(xlim = c(-4, 4), ylim = y_limits) +
    labs(title = custom_titles[nm]) +
    theme(
      panel.border = element_rect(
        colour = "black",
        fill = NA,
        linewidth = 0.6
      ),
      plot.title = element_text(
        size = 11,        # 👈 smaller titles
        face = "bold"
      ),
      axis.title.x = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      legend.text = element_text(size = 8),
      legend.title = element_text(size = 9),
      legend.key.size = unit(0.5, "lines")
    )
  
  return(p)
})

# ---- 4. Remove redundant axis titles ----
# Layout:
# [1] [2]
# [3] [4]

p1 <- plots[[3]] + theme(axis.title.x = element_blank())
p2 <- plots[[4]] + theme(axis.title.x = element_blank(),
                         axis.title.y = element_blank())
p3 <- plots[[1]] 
p4 <- plots[[2]] + theme(axis.title.y = element_blank())

# ---- 5. Combine with shared legend ----
library(patchwork)

combined_plot <- (
  (p1 + p2) /
  (p3 + p4)
) +
  plot_layout(guides = "collect") &
  theme(
    legend.position = "bottom",
    legend.box = "horizontal",
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 9),
    legend.key.size = unit(0.5, "lines")
  ) &
  guides(color = guide_legend(nrow = 2, byrow = TRUE),
         fill = guide_legend(nrow = 2, byrow = TRUE))

combined_plot
ggsave("Volcano_4panel_combined.png",
       combined_plot,
       width = 19,
        height = 14.85,
        units = "cm",
        dpi = 300)
```


## 10. Gene of interest
```{r}
library(dplyr)

# Create a summary of BCL2 across all kits
bcl2_summary <- lapply(names(dep_results), function(nm) {
  res <- dep_results[[nm]]
  
  # Make sure gene column exists
  gene_col <- if ("gene" %in% colnames(res)) "gene" else "name"
  
  res %>%
    filter(grepl("^map7$", !!sym(gene_col), ignore.case = TRUE)) %>%
    mutate(Kit = nm)
}) %>% bind_rows()

# View summary
bcl2_summary


# UniProt ID to check
uniprot_id <- "Q9UJW3"

# Search across all dep_results
uniprot_summary <- lapply(names(dep_results), function(nm) {
  res <- dep_results[[nm]]
  
  res %>%
    filter(grepl(uniprot_id, name, ignore.case = TRUE)) %>%
    mutate(Kit = nm)
}) %>% bind_rows()

# View result
uniprot_summary

uniprot_id <- "Q9Y6K1"

# Loop through all protein-level assays
for (nm in grep("_norm_protein$", names(qf), value = TRUE)) {
  protein_se <- qf[[nm]]
  
  found <- rowData(protein_se)$Proteins %>% 
           grepl(uniprot_id, ., ignore.case = TRUE)
  
  cat("Kit:", nm, "- P10415 present?", any(found), "\n")
  
  if (any(found)) {
    # Show the rows with this UniProt ID
    print(as.data.frame(rowData(protein_se)[found, ]))
  }
}
```

## 11. FC2FC common protein for VA
```{r}
kit_order <- c(
  "Kit4_2_norm_protein",  # Living cell short-term
  "Kit4_3_norm_protein",  # Living cell long-term
  "Kit4_4_norm_protein",  # Cell lysate short-term
  "Kit4_5_norm_protein"   # Expression long-term
)
```


```{r}
library(dplyr)

# Collect all VA log2FC values across all kits
all_fc <- lapply(kit_order, function(nm) {
  dep_results[[nm]]$VA_vs_C_ratio
}) %>% unlist()

lim <- max(abs(all_fc), na.rm = TRUE)
axis_limits <- c(-4, 4)
```

```{r}
# library(dplyr)
# library(ggplot2)
# library(viridis)

make_fc2fc_pair <- function(res1, res2, name1, name2,
                            expression_res = NULL) {
  
  df1 <- res1 %>%
    select(name,
           VA_vs_C_ratio,
           VA_vs_C_p.val) %>%
    rename(log2FC_1 = VA_vs_C_ratio,
           pval_1   = VA_vs_C_p.val)
  
  df2 <- res2 %>%
    select(name,
           VA_vs_C_ratio,
           VA_vs_C_p.val) %>%
    rename(log2FC_2 = VA_vs_C_ratio,
           pval_2   = VA_vs_C_p.val)
  
  merged <- inner_join(df1, df2, by = "name") %>%
    mutate(
      mean_negLogP = -log10((pval_1 + pval_2) / 2)
    )
  
  # Regression
  fit <- lm(log2FC_2 ~ log2FC_1, data = merged)
  slope <- coef(fit)[2]
  r2    <- summary(fit)$r.squared
  pval_reg <- summary(fit)$coefficients[2,4]
  
  label_text <- paste0(
    "Slope = ", round(slope, 3),
    "\nR² = ", round(r2, 3),
    "\nP = ", signif(pval_reg, 3)
  )
  
  p <- ggplot(merged, aes(log2FC_1, log2FC_2)) +
    
    geom_point(aes(fill = mean_negLogP),
               shape = 21,
               size = 1.6,
               alpha = 0.6,
               color = "gray40",  
               stroke = 0.6) +
    
    geom_smooth(method = "lm",
                se = FALSE,
                color = "red",   
                linewidth = 1.2) +
    
    annotate("text",
             x = axis_limits[1] * 0.9,
             y = axis_limits[2] * 0.9,
             label = label_text,
             hjust = 0,
             vjust = 1,
             size = 3,
             color = "blue") +
    
    scale_fill_viridis_c(
      option = "D",
      direction = -1,
      name = "-log10(p-value)"
    ) +
    
    coord_fixed(
      xlim = axis_limits,
      ylim = axis_limits,
      ratio = 1
    ) +
    
    theme_minimal(base_size = 11) +
    
    labs(
      title = paste(custom_titles[name1], "vs", custom_titles[name2]),
      x = paste0("log2FC (", custom_titles[name1], ")"),
      y = paste0("log2FC (", custom_titles[name2], ")")
    ) +
    
    theme(
      plot.title = element_text(face = "bold", size = 7),
      legend.position = "bottom",
      legend.text = element_text(size = 5),
      legend.title = element_text(size = 6),
      axis.title = element_text(size = 6)
    )
  
  return(p)
}
```




```{r}
p4 <- make_fc2fc_pair(dep_results[[kit_order[1]]],
                      dep_results[[kit_order[4]]],
                      kit_order[1], kit_order[4])

p5 <- make_fc2fc_pair(dep_results[[kit_order[3]]],
                      dep_results[[kit_order[4]]],
                      kit_order[3], kit_order[4])

p6 <- make_fc2fc_pair(dep_results[[kit_order[2]]],
                      dep_results[[kit_order[4]]],
                      kit_order[2], kit_order[4])

triple_plot_2 <- (p4 | p5 | p6) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

triple_plot_2

ggsave("FC2FC2Expression.png",
       triple_plot_2,
       width = 21,
        height = 10,
        units = "cm",
        dpi = 300)

```


```{r}
# library(ggplot2)
# library(dplyr)
# library(viridis)
# library(RColorBrewer)

make_fc2fc_pair1 <- function(res1, res2, res3, name1, name2, name3, axis_limits = c(-3, 3)) {
  
  # Merge the VA log2FC
  df1 <- res1 %>%
    select(name, VA_vs_C_ratio, gene) %>%
    rename(log2FC_1 = VA_vs_C_ratio)
  
  df2 <- res2 %>%
    select(name, VA_vs_C_ratio, gene) %>%
    rename(log2FC_2 = VA_vs_C_ratio)
  
  df3 <- res3 %>%
    select(name, VA_vs_C_ratio, gene) %>%
    rename(log2FC_3 = VA_vs_C_ratio)
  
  # Merge all three by 'name'
  merged <- df1 %>%
    full_join(df2, by = "name") %>%
    full_join(df3, by = "name")
  
  # Before labeling, create a gene column
merged <- merged %>%
  separate(gene, into = "gene", sep = ";", extra = "drop", fill = "right")
# Subset for labeling genes with large FC
label_df <<- merged %>%
  filter(abs(log2FC_1) > 1.5 | abs(log2FC_2) > 1.5)
  
  # Plot
  p <- ggplot(merged, aes(x = log2FC_1, y = log2FC_2)) +
    geom_point(aes(fill = log2FC_3),
               shape = 21, size = 1.5, alpha = 0.8, stroke = 0.5, color = "gray80") +
    
    # Add gene labels for large FC points
    ggrepel::geom_text_repel(
                    data = label_df,
                    aes(label = gene),   # use gene name instead of UniProt ID
                    size = 2,
                    box.padding = 0.3,
                    max.overlaps = 20,   # max number of labels allowed
                    force = 1, 
                    segment.color = "gray50"
                  ) +
    
    # Diverging color scale for FC of third kit
    scale_fill_gradient2(
      low = "blue",
      mid = "white",
      high = "red",
      midpoint = 0,
      limits = c(-0.3, 0.3),
      name = paste0("log2FC (", name3, ")"),
      oob = scales::squish
    ) +
    
    coord_fixed(xlim = axis_limits, ylim = axis_limits) +
    theme_minimal(base_size = 11) +
    labs(
      title = paste(name1, "vs", name2),
      x = paste0("log2FC (", name1, ")"),
      y = paste0("log2FC (", name2, ")")
    ) +
    theme(
      plot.title = element_text(face = "bold", size = 7),
      legend.position = "bottom",
      legend.text = element_text(size = 5),
      legend.title = element_text(size = 6),
      axis.title = element_text(size = 6)
    ) +
    guides(
      fill = guide_colorbar(barwidth = 3)
    )
  
  return(p)
}

```

```{r}
# First three plots (FC color)
p1 <- make_fc2fc_pair1(
  dep_results[["Kit4_3_norm_protein"]],
  dep_results[["Kit4_2_norm_protein"]],
  dep_results[["Kit4_5_norm_protein"]],
  "Long-term living cell CoPISA",
  "Short-term living cell CoPISA",
  "Long-term expression protoemics"
)


p2 <- make_fc2fc_pair1(
  dep_results[["Kit4_3_norm_protein"]],
  dep_results[["Kit4_4_norm_protein"]],
  dep_results[["Kit4_5_norm_protein"]],
  "Long-term living cell CoPISA",
  "Short-term Cell lysate CoPISA",
  "Long-term expression protoemics"
)


triple_plot_1 <- (p1 | p3) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

triple_plot_1

ggsave("FC2FCLongtoShort.png",
       triple_plot_1,
       width = 15,
        height = 10,
        units = "cm",
        dpi = 300)

```

## 12. Merge Four Proteomics Analyses by Fisher's Method for Combining P-values

```{r}
# -----------------------------
# 1. Prepare merged dataset
# -----------------------------

# Get common proteins across all datasets
common_proteins <- Reduce(intersect, list(
  dep_results$Kit4_2_norm_protein$gene,
  dep_results$Kit4_3_norm_protein$gene,
  dep_results$Kit4_4_norm_protein$gene,
  dep_results$Kit4_5_norm_protein$gene
))

# Filter datasets for common proteins
dep1 <- dep_results$Kit4_3_norm_protein %>% filter(gene %in% common_proteins) # "Long-term living cell CoPISA",
dep2 <- dep_results$Kit4_2_norm_protein %>% filter(gene %in% common_proteins) # "Short-term living cell CoPISA"
dep3 <- dep_results$Kit4_4_norm_protein %>% filter(gene %in% common_proteins) # "Short-term Cell lysate CoPISA"
dep4 <- dep_results$Kit4_5_norm_protein %>% filter(gene %in% common_proteins) # "Long-term expression proteomics"

deps <- list(dep1, dep2, dep3, dep4)

# Merge datasets by gene name and suffix columns with dataset index
merged <- imap(deps, function(df, i) {
  df %>%
    select(name = gene, ends_with("_p.val"), ends_with("_ratio")) %>%
    rename_with(~ paste0(.x, ".", i), -name)  # i is from imap
}) %>%
  purrr::reduce(inner_join, by = "name")

# -----------------------------
# 2. Helper functions
# -----------------------------

# Fisher method for combining p-values
combine_fisher <- function(...) {
  pvals <- c(...)
  pvals[is.na(pvals) | pvals == 0] <- 1
  chisq <- -2 * sum(log(pvals))
  df <- 2 * length(pvals)
  pchisq(chisq, df = df, lower.tail = FALSE)
}

# Harmonic mean for absolute fold change (pairwise)
hm2 <- function(x, y) {
  2 / (1/abs(x) + 1/abs(y))
}

# -----------------------------
# 3. Compute pairwise Fisher and HM fold changes
# -----------------------------

merged <- merged %>%
  rowwise() %>%
  mutate(
    # ---- A vs C ----
    Fisher_A_12 = combine_fisher(A_vs_C_p.val.1, A_vs_C_p.val.2),
    Fisher_A_13 = combine_fisher(A_vs_C_p.val.1, A_vs_C_p.val.3),
    Fisher_A_23 = combine_fisher(A_vs_C_p.val.2, A_vs_C_p.val.3),

    hmAbsFC_A_12 = hm2(A_vs_C_ratio.1, A_vs_C_ratio.2),
    hmAbsFC_A_13 = hm2(A_vs_C_ratio.1, A_vs_C_ratio.3),
    hmAbsFC_A_23 = hm2(A_vs_C_ratio.2, A_vs_C_ratio.3),

    # ---- V vs C ----
    Fisher_V_12 = combine_fisher(V_vs_C_p.val.1, V_vs_C_p.val.2),
    Fisher_V_13 = combine_fisher(V_vs_C_p.val.1, V_vs_C_p.val.3),
    Fisher_V_23 = combine_fisher(V_vs_C_p.val.2, V_vs_C_p.val.3),

    hmAbsFC_V_12 = hm2(V_vs_C_ratio.1, V_vs_C_ratio.2),
    hmAbsFC_V_13 = hm2(V_vs_C_ratio.1, V_vs_C_ratio.3),
    hmAbsFC_V_23 = hm2(V_vs_C_ratio.2, V_vs_C_ratio.3),

    # ---- VA vs C ----
    Fisher_VA_12 = combine_fisher(VA_vs_C_p.val.1, VA_vs_C_p.val.2),
    Fisher_VA_13 = combine_fisher(VA_vs_C_p.val.1, VA_vs_C_p.val.3),
    Fisher_VA_23 = combine_fisher(VA_vs_C_p.val.2, VA_vs_C_p.val.3),

    hmAbsFC_VA_12 = hm2(VA_vs_C_ratio.1, VA_vs_C_ratio.2),
    hmAbsFC_VA_13 = hm2(VA_vs_C_ratio.1, VA_vs_C_ratio.3),
    hmAbsFC_VA_23 = hm2(VA_vs_C_ratio.2, VA_vs_C_ratio.3)
  ) %>%
  ungroup()
# -----------------------------
# 4. Geometric mean cutoff for |log2FC|
# -----------------------------
# cutoffs_log2 <- c(0.3, 0.3)
# cutoffs_linear <- 2^cutoffs_log2
# gm_cutoff_linear <- exp(mean(log(cutoffs_linear)))
# gm_cutoff_log2 <- log2(gm_cutoff_linear)
gm_cutoff_log2 <- quantile(abs(plot_data$gmAbsFC), 0.75, na.rm = TRUE)
# -----------------------------
# 5. Prepare data for plotting
# -----------------------------
plot_data <- merged %>%
  mutate(
    # Convert HM |log2FC| to linear
    gmAbsFC_VA_12 = 2^hmAbsFC_VA_12,
    gmAbsFC_VA_13 = 2^hmAbsFC_VA_13,
    gmAbsFC_VA_23 = 2^hmAbsFC_VA_23
  ) %>%
  rowwise() %>%
  mutate(
    # Geometric mean in linear space (safeguard)
    gmAbsFC_VA_12 = exp(mean(log(c(gmAbsFC_VA_12)), na.rm = TRUE)),
    gmAbsFC_VA_13 = exp(mean(log(c(gmAbsFC_VA_13)), na.rm = TRUE)),
    gmAbsFC_VA_23 = exp(mean(log(c(gmAbsFC_VA_23)), na.rm = TRUE))
  ) %>%
  ungroup() %>%
  mutate(
    # Back to log2 for plotting
    gmAbsFC_VA_12 = log2(gmAbsFC_VA_12),
    gmAbsFC_VA_13 = log2(gmAbsFC_VA_13),
    gmAbsFC_VA_23 = log2(gmAbsFC_VA_23)
  ) %>%
  select(
    name,
    gmAbsFC_VA_12, Fisher_VA_12,
    gmAbsFC_VA_13, Fisher_VA_13,
    gmAbsFC_VA_23, Fisher_VA_23
  ) %>%
  pivot_longer(
    cols = -name,
    names_to = c(".value", "Condition"),
    names_pattern = "(gmAbsFC|Fisher)_(VA_\\d+)"
  ) %>%
  mutate(
    Condition = recode(Condition,
      "VA_12" = "Long-term living cell CoPISA vs \nShort-term living cell CoPISA",
      "VA_13" = "Long-term living cell CoPISA vs \nShort-term Cell lysate CoPISA",
      "VA_23" = "Short-term living cell CoPISA vs \nShort-term Cell lysate CoPISA"
    ),
    significant = ifelse(Fisher < 0.05 & gmAbsFC > gm_cutoff_log2,
                         "Significant", "Not significant")
  )

# -----------------------------
# 6. Volcano-style plot
# -----------------------------
library(ggplot2)
library(ggrepel)
# --- 1. Remove multiple gene names and keep the first one to show 
plot_data <- plot_data %>%
  mutate(name_plot = sub(";.*$", "", name))  # take everything before the first semicolon

# --- 2. Add dep4 fold-change to plot_data ---
dep4_fc <- dep_results$Kit4_5_norm_protein %>%
  filter(gene %in% common_proteins) %>%
  select(name = gene, FC_dep4 = VA_vs_C_ratio)  # or whichever comparison you want

plot_data <- plot_data %>%
  left_join(dep4_fc, by = c("name" = "name")) %>%
  mutate(
    # keep only first gene if multiple genes
    name_plot = sub(";.*$", "", name)
  )

# Volcano-style plot with cutoff lines
spray <- ggplot(plot_data, aes(x = gmAbsFC, y = -log10(Fisher))) +
  geom_point(aes(
    color = ifelse(abs(gmAbsFC) > gm_cutoff_log2 & Fisher < 0.05,
                   dep_results$Kit4_5_norm_protein$VA_vs_C_ratio[match(name, dep_results$Kit4_5_norm_protein$gene)],
                   NA)
  ), alpha = 0.8, size = 2) +
  scale_color_gradient2(
    low = "blue", mid = "gray90", high = "red",
    midpoint = 0,
    na.value = "gray70",
    name = "Long-term expression proteomics log2FC"
  ) +
  geom_text_repel(
    data = subset(plot_data, abs(gmAbsFC) > gm_cutoff_log2 & Fisher < 0.05),
    aes(label = sub(";.*$", "", name)),
    size = 3,
    max.overlaps = 20,
    box.padding = 0.4,
    point.padding = 0.3,
    segment.color = "grey50",
    show.legend = FALSE
  ) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "blue") +
  geom_vline(xintercept = gm_cutoff_log2, linetype = "dashed", color = "blue") +
  facet_wrap(~Condition, scales = "free_y", ncol = 1) +
  theme_bw(base_size = 13) +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(fill = "white", color = "black"),
    panel.grid = element_blank()
  ) +
  labs(x = "Geometric Mean |log2 Fold Change|",
       y = "-log10(Fisher Combined p-value)"
  )

spray
# Save your plot 'spray' as PNG
# ggsave(
#   filename = "sprayPlot.png",
#   plot = spray,          # your ggplot object
#   width = 17,            # width in cm
#   height = 9,            # height in cm
#   units = "cm",
#   dpi = 300
# )

```

# 13. R1-R4 group of proteins
```{r}
library(dplyr)

short_term_subset <- plot_data %>%
  filter(
    Condition == "Short-term living cell CoPISA vs \nShort-term Cell lysate CoPISA"
  )

short_term_subset <- short_term_subset %>%
  mutate(
    absent_in_long_term = !name %in% (
      plot_data %>%
        filter(
          Condition %in% c(
            "Long-term living cell CoPISA vs \nShort-term Cell lysate CoPISA",
            "Long-term living cell CoPISA vs \nShort-term living cell CoPISA"
          ),
          significant == "Significant"
        ) %>%
        pull(name)
    )
  )

table(short_term_subset$significant, short_term_subset$absent_in_long_term)

short_term_subset <- short_term_subset %>%
  mutate(
    state = case_when(
      significant == "Significant" & !absent_in_long_term ~ "Sig_in_long_term",
      significant == "Significant" & absent_in_long_term  ~ "Sig_not_in_long_term",
      significant == "Not significant" & !absent_in_long_term ~ "NS_in_long_term",
      TRUE ~ "NS_not_in_long_term"
    )
  )

genes_by_state <- short_term_subset %>%
  group_by(state) %>%
  summarise(
    genes = list(name),
    fc_values = list(FC_dep4),
    .groups = "drop"
  )

# Drug MoA in Resistant cells
(sig_long_term <- short_term_subset %>%
  filter(state == "Sig_in_long_term") %>%
  select(name, FC_dep4)%>% 
  arrange(desc(FC_dep4)))

# Drug MoA in Sensitive cells
(sig_not_long_term <- short_term_subset %>%
  filter(state == "Sig_not_in_long_term") %>%
  select(name, FC_dep4) %>% 
  arrange(desc(FC_dep4)))

# (ns_long_term <- short_term_subset %>%
#   filter(state == "NS_in_long_term") %>%
#   select(name, FC_dep4)%>% 
#   arrange(desc(FC_dep4)))
# 
# (ns_not_long_term <- short_term_subset %>%
#   filter(state == "NS_not_in_long_term") %>%
#   select(name, FC_dep4)%>% 
#   arrange(desc(FC_dep4)))
```

```{r}
library(dplyr)
library(ggplot2)
library(ggrepel)

df1 <- sig_long_term %>%
  arrange(FC_dep4) %>%
  mutate(
    gene = factor(name, levels = name),  # preserves ranking
    group = ifelse(FC_dep4 > 0.5, "> 0.5", "≤ 0.5")
  )

df2 <- sig_not_long_term %>%
  arrange(FC_dep4) %>%
  mutate(
    gene = factor(name, levels = name),  # preserves ranking
    group = ifelse(FC_dep4 > 0.5, "> 0.5", "≤ 0.5")
  )



r1_2 <- ggplot(df1, aes(x = gene, y = FC_dep4)) +
  
  # stems
  geom_segment(aes(xend = gene, y = 0, yend = FC_dep4, color = group),
               linewidth = 0.6) +
  
  # zero reference line
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  
  # bubbles
  geom_point(aes(color = group), size = 3) +
  
  scale_color_manual(values = c("> 0.5" = "#EA4335", "≤ 0.5" = "#4285F4")) +
  
  coord_flip() +   # 🔥 flip AFTER ordering
  
  theme_minimal(base_size = 12) +
  
  theme(
    axis.text.y = element_text(size = 10),  # gene labels now on y-axis
    axis.ticks = element_blank(),
    
    panel.grid = element_blank(),
    
    panel.border = element_rect(
      colour = "black",
      fill = NA,
      linewidth = 0.8
    ),
    
    legend.position = "bottom"
  ) +
  
  labs(
    x = NULL,
    y = NULL,
    color = "Long-term expression proteomics log2FC",
    title = "Long-term \nretained targets"
  )

r3_4 <- ggplot(df2, aes(x = gene, y = FC_dep4)) +
  
  # stems
  geom_segment(aes(xend = gene, y = 0, yend = FC_dep4, color = group),
               linewidth = 0.6) +
  
  # zero reference line
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  
  # bubbles
  geom_point(aes(color = group), size = 3) +
  
  scale_color_manual(values = c("> 0.5" = "#EA4335", "≤ 0.5" = "#4285F4")) +
  
  coord_flip() +   # 🔥 flip AFTER ordering
  
  theme_minimal(base_size = 12) +
  
  theme(
    axis.text.y = element_text(size = 10),  # gene labels now on y-axis
    axis.ticks = element_blank(),
    
    panel.grid = element_blank(),
    
    panel.border = element_rect(
      colour = "black",
      fill = NA,
      linewidth = 0.8
    ),
    
    legend.position = "bottom"
  ) +
  
  labs(
    x = NULL,
    y = NULL,
    color = "Long-term expression proteomics log2FC",
    title = "Long-term \nlost targets"
  )
```


```{r}
library(patchwork)

combined_plot <- (
  (r3_4  + r1_2) ) +
  plot_layout(guides = "collect") &
  theme(
    plot.title = element_text(face = "bold"),
    legend.position = "bottom",
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 9),
    legend.key.size = unit(0.6, "lines")
  )

combined_plot
```

```{r}
library(tidyverse)
library(tidyr)
# install.packages("devtools")
# devtools::install_github("ricardo-bion/ggradar")
library(ggradar)
library(scales)
library(scales)
library(gridExtra)

dep1_fc <- dep_results$Kit4_3_norm_protein %>%
  filter(gene %in% common_proteins) %>%
  select(name = gene, `Long-term living cell CoPISA` = VA_vs_C_ratio)

dep2_fc <- dep_results$Kit4_2_norm_protein %>% 
  filter(gene %in% common_proteins) %>%
  select(name = gene, `Short-term living cell CoPISA` = VA_vs_C_ratio)

dep3_fc <- dep_results$Kit4_4_norm_protein %>% 
  filter(gene %in% common_proteins) %>%
  select(name = gene, `Short-term Cell lysate CoPISA` = VA_vs_C_ratio)

dep4_fc <- dep_results$Kit4_5_norm_protein %>% 
  filter(gene %in% common_proteins) %>%
  select(name = gene, `Long-term expression proteomics` = VA_vs_C_ratio)

# Merge datasets
df_wide <- purrr::reduce(
  list(dep1_fc, dep2_fc, dep3_fc, dep4_fc),
  full_join,
  by = "name"
)

selected_proteins <- c("NRP2", "RPL18", "PLP2", "RPS28", "NOTCH1")

# Prepare radar data
df_radar <- df_wide %>%
  filter(name %in% selected_proteins) %>%
  replace(is.na(.), 0) %>%
  mutate(across(
    -name,
    ~ rescale(.x, to = c(0, 1), from = c(-3, 3))
  ))

# Create one radar plot per gene
plot_list <- lapply(1:nrow(df_radar), function(i) {

  gene_df <- df_radar[i, ]

  ggradar(
    gene_df,
    background.circle.colour = "white",
    gridline.mid.colour = "gray80",
    gridline.max.colour = "gray70",
    gridline.min.colour = "gray90",
    axis.line.colour = "gray60",
    group.line.width = 1.5,
    group.point.size = 3,
    group.colours = "steelblue",
    legend.position = "none",
    values.radar = c("0", "0.5", "1")
  ) +
    ggtitle(gene_df$name)
})

# Arrange plots
grid.arrange(grobs = plot_list, ncol = 2)
```


# 14. Combine mzMl and identification
```{r}
library(Spectra)
files <- c("/Users/nvd448/Documents/VenAza Project/mzML/001_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/002_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/003_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/004_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/005_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/006_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/007_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/008_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/009_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/010_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/011_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/012_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/013_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/014_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/015_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/016_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/017_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/018_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/019_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/020_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/021_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/022_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/023_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/024_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/025_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/026_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/027_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/028_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/029_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/030_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/031_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/032_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/033_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/034_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/035_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/036_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/037_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/038_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/039_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/040_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/041_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/042_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/043_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/044_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/045_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/046_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/047_4TMT16_2DTMT_G24.mzML",
  "/Users/nvd448/Documents/VenAza Project/mzML/048_4TMT16_2DTMT_G24.mzML")

allKit_mzML <- Spectra(files, source = MsBackendMzR(), backendInitialize = TRUE)
spectraVariables(allKit_mzML)

# # Extract numeric file index
# allKit_mzML$file_index <- as.integer(
#     sub("_.*", "", allKit_mzML$file)
# )

```


```{r}
# Mapping to MaxQuant results

library(data.table)

mq <- fread("~/Documents/VenAza Project/msms.txt")
colnames(mq)


# Create clean file names in Spectra
allKit_mzML$file <- basename(allKit_mzML$dataOrigin)
allKit_mzML$file <- sub("\\.mzML$", "", allKit_mzML$file)

# Extract numeric file index
allKit_mzML$file_index <- as.integer(
    sub("_.*", "", allKit_mzML$file)
)

# Define treatment groups
allKit_mzML$treatment <- cut(
    allKit_mzML$file_index,
    breaks = c(0, 12, 24, 36, 48),
    labels = c(
        "Short-term living cell CoPISA",
        "Long-term living cell CoPISA",
        "Short-term Cell lysate CoPISA",
        "Long-term expression proteomics"
    )
)

table(allKit_mzML$treatment)

# Create clean file names in mq
mq$file <- mq$`Raw file`
mq$file <- sub("\\.raw$", "", mq$file, ignore.case = TRUE)

# Build matching key in Spectra and mq
allKit_mzML$key <- paste(
    allKit_mzML$file,
    allKit_mzML$acquisitionNum,
    sep = "_"
)

mq$key <- paste(
    mq$file,
    mq$`Scan number`,
    sep = "_"
)

# Map and combine
spid <- match(allKit_mzML$key, mq$key)

allKit_mzML$sequence <- mq$Sequence[spid]
allKit_mzML$modified_sequence <- mq$`Modified sequence`[spid]
allKit_mzML$score <- mq$Score[spid]
allKit_mzML$pep <- mq$PEP[spid]
allKit_mzML$proteins <- mq$Proteins[spid]
```

# 15. Protein-centric workflow

```{r}
# =========================================================
# COMPLETE PUBLICATION-STYLE MS/MS PIPELINE
# Protein -> Best peptide -> Best spectra -> Fragmentation plot
# =========================================================

library(dplyr)
library(data.table)
library(reshape2)
library(ggplot2)
library(PSMatch)
library(patchwork)
library(cowplot)
library(gridExtra)

# =========================================================
# 1. FILTER IDENTIFIED MS2 SPECTRA
# =========================================================

id_spec <- allKit_mzML[
    !is.na(allKit_mzML$sequence) &
    allKit_mzML$msLevel == 2
]

# =========================================================
# 2. SELECT PROTEIN OF INTEREST
# =========================================================

protein_id <- "O60462"   # NRP2
Gene_name <- "NRP2"
prot_spec <- id_spec[
    grepl(protein_id, id_spec$proteins)
]

cat("Number of spectra from protein:", length(prot_spec), "\n")

# =========================================================
# 3. PEPTIDE SUMMARY + RANKING
# =========================================================

dt <- as.data.table(
    spectraData(prot_spec)
)

# peptide statistics per treatment
pep_summary <- dt[, .(
    n_spectra = .N,
    mean_score = mean(score, na.rm = TRUE),
    median_score = median(score, na.rm = TRUE),
    mean_pep = mean(pep, na.rm = TRUE),
    median_pep = median(pep, na.rm = TRUE)
), by = .(sequence, treatment)]

# peptide statistics globally
pep_summary_all <- dt[, .(
    n_spectra = .N,
    mean_score = mean(score, na.rm = TRUE),
    median_pep = median(pep, na.rm = TRUE),
    n_treatments = uniqueN(treatment)
), by = sequence]

# consistency metric
pep_summary_all[, consistency :=
    n_treatments / max(n_treatments)
]

# final ranking score
pep_summary_all[, final_score :=
    mean_score *
    consistency *
    log1p(n_spectra)
]

# rank peptides
pep_ranked <- pep_summary_all[
    order(-final_score)
]

top_peptide <- pep_ranked$sequence[1]

cat("Selected peptide:", top_peptide, "\n")

# =========================================================
# 4. SELECT BEST SPECTRA PER TREATMENT
# =========================================================

pep_spec <- id_spec[
    id_spec$sequence == top_peptide
]

spec_list <- split(
    pep_spec,
    pep_spec$treatment
)

# best spectrum index per treatment
best_idx <- tapply(
    seq_along(pep_spec$score),
    pep_spec$treatment,
    function(i) i[
        which.max(pep_spec$score[i])
    ]
)

best_spec <- pep_spec[
    unlist(best_idx)
]

# =========================================================
# 5. GENERATE THEORETICAL FRAGMENTS
# =========================================================

frags <- calculateFragments(top_peptide)

frag_df <- data.frame(frags) %>%
    filter(mz > 0)

frag_df$ion_type <- ifelse(
    grepl("^b", frag_df$ion),
    "b",
    "y"
)

# =========================================================
# 6. EXTRACT MS2 PEAKS
# =========================================================

best_list <- split(
    best_spec,
    best_spec$treatment
)

extract_ms2 <- function(spec, treatment_name) {

    data.frame(
        mz = as.numeric(mz(spec)[[1]]),
        intensity = as.numeric(intensity(spec)[[1]]),
        treatment = treatment_name
    )
}

ms2_df <- do.call(
    rbind,
    lapply(names(best_list), function(tr) {

        extract_ms2(
            best_list[[tr]][1],
            tr
        )
    })
)

# =========================================================
# 7. SANITY CHECKS
# =========================================================

stopifnot(
    "mz" %in% names(ms2_df),
    "intensity" %in% names(ms2_df),
    is.numeric(ms2_df$mz),
    is.numeric(ms2_df$intensity)
)

# =========================================================
# 8. NORMALIZE INTENSITIES
# =========================================================

ms2_df <- ms2_df %>%
    group_by(treatment) %>%
    mutate(
        intensity =
            intensity /
            max(intensity, na.rm = TRUE)
    ) %>%
    ungroup()

# =========================================================
# 9. MATCH FRAGMENTS TO PEAKS
# =========================================================

match_fragments <- function(ms2, frag, tol = 0.5) {

    frag$peak_mz <- NA
    frag$peak_intensity <- NA
    frag$matched <- FALSE

    for (i in seq_len(nrow(frag))) {

        idx <- which(
            abs(ms2$mz - frag$mz[i]) <= tol
        )

        if(length(idx) > 0) {

            best <- idx[
                which.max(ms2$intensity[idx])
            ]

            frag$peak_mz[i] <- ms2$mz[best]
            frag$peak_intensity[i] <- ms2$intensity[best]
            frag$matched[i] <- TRUE
        }
    }

    frag
}

annotated_list <- lapply(
    split(ms2_df, ms2_df$treatment),
    function(df) {
        match_fragments(
            df,
            frag_df
        )
    }
)

frag_annot <- bind_rows(
    annotated_list,
    .id = "treatment"
)

# =========================================================
# 10. PLOT FUNCTION (POLISHED)
# =========================================================

plot_spectrum <- function(ms2, annot, panel_title) {

    ggplot(ms2, aes(x = mz, y = intensity)) +

        # raw spectrum
        geom_segment(
            aes(xend = mz, yend = 0),
            linewidth = 0.25,
            color = "black"
        ) +

        # b ions
        geom_segment(
            data = subset(annot, ion_type == "b" & matched),
            aes(x = peak_mz, xend = peak_mz, y = 0, yend = peak_intensity),
            color = "royalblue",
            linewidth = 0.5
        ) +

        # y ions
        geom_segment(
            data = subset(annot, ion_type == "y" & matched),
            aes(x = peak_mz, xend = peak_mz, y = 0, yend = peak_intensity),
            color = "forestgreen",
            linewidth = 0.5
        ) +

        # matched peaks
        geom_point(
            data = subset(annot, matched),
            aes(x = peak_mz, y = peak_intensity),
            color = "red",
            size = 1.6
        ) +

        # ion labels
        geom_text(
            data = subset(annot, matched),
            aes(x = peak_mz, y = peak_intensity, label = ion, color = ion_type),
            size = 3,
            fontface = "bold",
            vjust = -0.7,
            show.legend = FALSE
        ) +

        scale_color_manual(values = c(b = "royalblue", y = "forestgreen")) +

        labs(
            title = NULL,
            x = "m/z",
            y = "Relative intensity"
        ) +

        ylim(0, 1.15) +

        # ✔ single clean metadata tag (corner label)
        annotate(
            "label",
            x = Inf,
            y = Inf,
            label = panel_title,
            hjust = 1.05,
            vjust = 1.2,
            size = 3.3,
            fontface = "bold",
            label.size = 0.25
        ) +

        theme_classic(base_size = 12) +

        theme(
            axis.title = element_text(face = "bold")
        )
}

# =========================================================
# 11. BUILD SPECTRUM PANELS (POLISHED)
# =========================================================

panel_list <- lapply(
    unique(ms2_df$treatment),
    function(tr) {

        ms2_sub <- subset(ms2_df, treatment == tr)
        ann_sub <- subset(frag_annot, treatment == tr)

        # extract metadata safely
        score_val <- best_spec$score[best_spec$treatment == tr][1]
        charge <- best_spec$precursorCharge[best_spec$treatment == tr][1]

        panel_title <- paste0(
            tr,
            " | score: ", round(score_val, 2),
            " | z=", charge, "+"
        )

        plot_spectrum(ms2_sub, ann_sub, panel_title)
    }
)

main_panels <- wrap_plots(panel_list, ncol = 2)

# =========================================================
# 12. PEPTIDE COVERAGE MAP
# =========================================================

seq_df <- data.frame(
    aa = strsplit(top_peptide, "")[[1]],
    pos = seq_along(
        strsplit(top_peptide, "")[[1]]
    )
)

top_map <- ggplot(
    seq_df,
    aes(
        x = pos,
        y = 0
    )
) +

    geom_text(
        aes(label = aa),
        size = 6,
        fontface = "bold"
    ) +

    geom_segment(
        aes(
            x = pos - 0.45,
            xend = pos + 0.45,
            y = 0.1,
            yend = 0.1
        ),
        color = "royalblue",
        linewidth = 1.2
    ) +

    geom_segment(
        aes(
            x = pos - 0.45,
            xend = pos + 0.45,
            y = -0.1,
            yend = -0.1
        ),
        color = "forestgreen",
        linewidth = 1.2
    ) +

    annotate(
        "text",
        x = 1,
        y = 0.9,
        label = "b ions",
        color = "royalblue",
        fontface = "bold"
    ) +

    annotate(
        "text",
        x = 1,
        y = -0.9,
        label = "y ions",
        color = "forestgreen",
        fontface = "bold"
    ) +

    theme_void()

# =========================================================
# 13. METADATA PANEL
# =========================================================

precursor_mz <- round(
    precursorMz(best_spec)[1],
    4
)

charge <- best_spec$precursorCharge[1]

meta_text <- paste0(
    "Protein: ", protein_id, "\n",
    "Gene name: ", Gene_name, "\n",
    #"Peptide: ", top_peptide, "\n",
    "Precursor m/z: ",
    precursor_mz,
    " (",
    charge,
    "+)\n",
    "Peptide length: ",
    nchar(top_peptide),
    " aa\n",
    "Fragment ion tolerance: ±0.5 Da"
)

meta_plot <- ggplot() +

    annotate(
        "text",
        x = 0,
        y = 1,
        label = meta_text,
        hjust = 0,
        vjust = 1,
        size = 4
    ) +

    theme_void()

# =========================================================
# 14. FINAL LAYOUT
# =========================================================

final_plot <- plot_grid(
    #meta_plot,
    #top_map,
    main_panels,
    ncol = 1,
    rel_heights = c(
        0.12,
        0.15,
        0.73
    )
)

# =========================================================
# 15. TITLE
# =========================================================

# title_plot <- ggdraw() +
# 
#     draw_label(
#         paste0(
#             "MS/MS Fragmentation of Peptide: ",
#             top_peptide
#         ),
#         fontface = "bold",
#         size = 20
#     )

# =========================================================
# 16. COMPLETE FIGURE
# =========================================================

publication_figure <- plot_grid(
    #title_plot,
    final_plot,
    ncol = 1,
    rel_heights = c(
        0.06,
        0.94
    )
)

publication_figure


```
